### 2022-01-28 appendix_math_background/appendix.tex

%!TEX root = ../thesis.tex
\chapter{More mathematical background}\label{app: math background}
\section{Time-integration schemes}\label{sec: time integration}
% SSPRK
% CFL