### 2022-01-28 chapter5/chapter.tex

%!TEX root = ../thesis.tex
\chapter{Global convergence results}\label{ch: global results}
In this chapter we prove some results which stem from the main results of Chapter~\ref{sec: conv}. We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process at the of points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\). In Corollary~\ref{cor: aln22222212} we prove that the transition probability of the embedded process of the QBD-RAP converge those of the embedded process of the fluid queue. In Corollary~\ref{cor: short cor} we prove that distribution of the sojourn time of the QBD-RAP in a given level converges to distribution of the sojourn time of the fluid queue in the corresponding interval. In Section~\ref{sec: sub global}, we state global results on the weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.

In this chapter we work with the augmented state space scheme to model phases with rates \(c_i=0\) as described in Section~\ref{sec: zero states}. The results of this chapter rely on results of Chapter~\ref{sec: conv} and therefore apply to the QBD-RAP scheme which uses ephemeral states to model a fluid queue which starts in a phase with rate \(0\), as described in Section~\ref{sec: initial conditions}. However, supplementing the results of Chapter~\ref{sec: conv} with the results from Appendix~\ref{app:extend conv} and using the same arguments from this chapter with only slight modifications, then the results of this chapter can be extended to the augmented state-space QBD-RAP scheme without the initial ephemeral phases.

% In this chapter we prove a global convergence result. We partition the Laplace tarnsforms with respect to time of the QBD-RAP on the number of level changes. Then, by considering the process at changes of level, we are able to work with the discrete-time Markov chain embedded at changes of level. First, we show that the transition probabilities of the embedded chain of the QBD-RAP convergence to the transition probabilities of the chain embedded in the fluid queue. We show that for a fixed number of level changes, \(n\), say, the embedded process of the QBD-RAP converges to an embedded discrete time process of the fluid queue, embedded at the times when the fluid queue leaves the bands \(\mathcal D_{\ell}\). Now, from Chapter~\ref{sec: conv} we have a result on the convergence of the QBD-RAP up to the first change of level (THM) and at the first change of level (COR). By time-homogeneity and the strong Markov property of the QBD-RAP then a Corollary to (COR) is the convergence of the transition probabilities of the embedded process of the QBD-RAP to the transition probabilities of the embedded process of the fluid queue. Similarly, a Corollary to (THM) is the convegence of the QBD-RAP to the fluid queue between the \(n\)th and \(n+1\)th change of level. Putting these Corollaries together, we can than claim a convergence of the distribution of QBD-RAP to the distribution of the fluid queue between the \(n\)th and \(n+1\)th change of level. Next, by summing over the partition on the number of level changes, \(n\), and applying the Dominated Convergence Theorem, we can then claim that Laplace transform with respect to time of the fluid queue converges to the Laplace transform with respect to time of the fluid queue. Finally, by the Extended Continuity Theorem, we can then claim that the distributions of the QBD-RAP and fluid queue converge for almost all \(t\geq 0\).

% \begin{enumerate}
% 	\item Define the embedded process at level changes of the QBD-RAP and the correspoding embedded process of the fluid queue, then use the results of Chapter~\ref{sec: conv} to claim convergence of the embedded processes; that is, convergence at the \(n\)th level change, \(n\geq 1\). 
% 	\item Extend the convergence of the embedded processes to convergence of the Laplace transforms between the \(n\)th and \(n+1\)th change of level, \(n\geq 0\).
% 	\item Apply the Dominated Convergence Theorem to claim that the Laplace transform with respect to time of the QBD-RAP converges to the Laplace transform with respect to time of the fluid queue. 
% 	\item Apply the Extended Continuity Theorem for Laplace transforms to claim the the distributions of the processes converge for almost all \(t\geq 0\). 
% \end{enumerate}

% The structure of this chapter is as follows. We start by showing weak convergence of the approximation at the \(1\)st and then \(n\)th change of level, \(n> 1\) in Sections~\ref{sec: 1st change} and \ref{sec: nth change}. Section~\ref{sec: between n and np1} generalises this to show weak convergence between the \(n\)th and \(n+1\)th change of level. Via the Dominated Convergence Theorem (again), Section~\ref{sec: local to global} proves the weak convergence of the QBD-RAP approximation to the fluid queue. So as to not obscure the logical flow of the presentation, technical results are reserved for the Appendices.

\section{Convergence of an embedded process}\label{sec: embedded}
% Before we move on to global convergence we need another result about convergence of the QBD-RAP scheme and the fluid queue at changes of level.

% Regarding Remark~\ref{rem: point wies}, we do require convergence at certain boundary points which correspond to changes of level of the QBD-RAP. To this end, we have the following result.

In this section we consider the embedded process formed by observing the QBD-RAP at the orbit restart epochs. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
\[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
be orbit restart epochs. These are the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary. To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed. Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\). The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\). The process \(\{\Ydp(n)\}_{n\geq 0}\) is a discrete-time Markov chain, which is time-homogeneous for \(n\geq 1\).

Let \(\{\tau_n^X\}_{n\geq 0}\), be the sequence of (stopping) times with \(\tau_0^X=0\), and 
\[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\},\]
for \(n\geq 0\). For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or leaves a boundary, for the \(n\)th time. The embedded process \(\{\bs X(\tau_n)\}\) is a discrete-time Markov chain which is time-homogeneous for \(n\geq 1\).

We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we will utilise later to prove a global result.
\begin{cor}\label{cor: aln222222} For \(\ell_0\in\mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
	for \(n=0\), then
	\begin{align}
		&\mathbb P(\Ydp(1) = (\ell(\ell_0,j),j), \tau_1^{(p)}\leq E^\lambda
            	 \mid \bs Y^{(p)}(0) = (\ell_0,\bs a_{\ell_0,i}^{(p)}(x_0),i)) \nonumber
	 	%
		\\&\to  
			\mathbb P(\bs X(\tau_1^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_1^{X}\leq E^\lambda \mid \bs X(0) = (x_0,i)),\label{eqn: lim 1}
	\end{align}
	and for \(n\geq 1\), then, 
	\begin{align}
		&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j), \tau_{n+1}^{(p)}\leq E^\lambda
            	 \mid \Ydp(n) = (\ell_0,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
	 	%
		\\&\to  
			\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd}
	\end{align}
	where \(\ell(\ell_0,j)\) can take values
	\begin{align}
		\ell(\ell_0,j) = \begin{cases}
			\ell_0-1, & \ell_0\in\{0,1,\dots,K+1\},\, j\in\calS_-\\
			%
			\ell_0, & \ell_0 = 0, j\in\mathcal S_+, \mbox{ or }\ell_0 = K, j\in\mathcal S_-,\\
			%
			\ell_0+1, & \ell_0\in\{-1,0,1,\dots,K\},\, j\in\calS_+.
		\end{cases}
	\end{align}
	% and 
	% \begin{align}
	% 	y(\ell,\ell_0,j) = \begin{cases}
	% 		y_{\ell_0}, & \ell=\ell_0-1,\, j\in\calS_-, \mbox{ or } \ell=\ell_0=0,\, j \in\calS_+, \\
	% 		y_{\ell_0+1}, & \ell=\ell_0+1,\, j\in\calS_+, \mbox{ or } \ell=\ell_0=K,\, j \in\calS_-.
	% 	\end{cases}
	% \end{align}
\end{cor}
\begin{proof}
The convergence for \(\ell_0\in\{-1,K+1\}\) holds trivially as the QBD-RAP and fluid queue have the same behaviour at the sticky boundary.

The case for \(n=0\) is a direct result of Corollary~\ref{cor: aln222}. % For the case with \(n=0\), since \(\tau_1^{(p)}<\infty\) almost surely for all \(p\geq 1\), as is \(\tau_1^X\), then upon taking \(\lambda \to 0\) in the results in Corollary~\ref{cor: aln222} we get the convergence in (\ref{eqn: lim 1}).

All that is left is to prove (\ref{eqn: ddddddd}). For \(n\geq 1\), consider the transition probabilities of the embedded process from the QBD-RAP, 
\begin{align}
	&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j),\tau_{n+1}^{(p)}\leq E^\lambda
	\mid \Ydp(n) = (\ell_0,i),\tau_{n}^{(p)}\leq E^\lambda) \nonumber 
	\\&=\mathbb P(\Ydp(1) = (\ell(\ell_0,j),j), \tau_{1}^{(p)}\leq E^\lambda
	\mid \bs Y^{(p)}(0)=(\ell_0 ,\bs\alpha, i)), \label{eqn: aaannnn}
\end{align}
since the QBD-RAP is time-homogeneous and the exponential random variable \(E^\lambda\) is memoryless. %The probability on the right of (\ref{eqn: aaannnn}) is equal to
% \begin{align}
% 	&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1(j\in\calS_-)+1(j\in\calS_+), \varphi(\tau_1^{(p)}) = j
% 			 \mid \bs Y^{(p)}(0) = (\ell_0 ,\bs\alpha, i)).\nonumber
% 	%  \intertext{for \(j\in\calS_+\)}
% 	%  &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j
% 	% 		 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
% \end{align}
% Thus, once again noting that \(\tau_1^{(p)}<\infty\) almost surely, then taking \(\lambda \to 0\) in 
Applying Corollary~\ref{cor: aln222} to (\ref{eqn: aaannnn}) then
\begin{align}
	&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j),\tau_{n+1}^{(p)}\leq E^\lambda
	\mid \Ydp(n) = (\ell_0,i),\tau_{n}^{(p)}\leq E^\lambda) \nonumber 
	\\&\to \mathbb P(\bs X(\tau_{1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(0) = (y_{\ell_0+1(i\in\calS_-)},i)).\label{eqnL sdfdfaaa}
\end{align}
Since the fluid queue is time-homogeneous, and \(E^\lambda\) memoryless, then (\ref{eqnL sdfdfaaa}) is equal to 
\begin{align}
	&\mathbb P(\bs X(\tau_{1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(\tau_{1}^X) = (y_{\ell_0+1(i\in\calS_-)},i)) \nonumber
	\\&=\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_{n}^{X}\leq E^\lambda ),
\end{align}
which proves the result. 
\end{proof}
A direct corollary of Corollary~\ref{cor: aln222222} is the convergence of the transition probabilities of the embedded process. 
\begin{cor}\label{cor: aln22222212} For \(\ell_0\in \mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
		for \(n=0\), then
		\begin{align}
			&\mathbb P(\Ydp(1) = (\ell(\ell_0,j),j)
					 \mid \bs Y^{(p)}(0) = (\ell_0,\bs a_{\ell_0,i}^{(p)}(x_0),i)) \nonumber
			 %
			\\&\to  
				\mathbb P(\bs X(\tau_1^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j)  \mid \bs X(0) = (x_0,i)).\label{eqn: lim 2}
		\end{align}
		and for \(n\geq 1\),  
		\begin{align}
			&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j) 
					 \mid \Ydp(n) = (\ell_0,i) ) \nonumber
			 %
			\\&\to  
				\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j) \mid \bs X(\tau_{n}) = (y_{\ell_0+1(i\in\calS_-)},i) ).\label{eqn: ddddddd2}
		\end{align}
\end{cor}
\begin{proof}
	Since \(\tau_1^{(p)}<\infty\) almost surely, as is \(\tau_1^X\), then taking \(\lambda \to 0\) in Corollary~\ref{cor: aln222222} yields the result. 
\end{proof}
Corollary~\ref{cor: aln22222212} states that the transition probabilities of the embedded processes converge. Thus, the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the sequence of distributions of \(\{\Ydp(n)\}\) is tight. Thus, we can establish the convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).

Another direct corollary of Corollary~\ref{cor: aln222} is the convergence in distribution of the random variables \(\{\tau_1^{(p)}\}_{p\geq 1}\) to \(\tau_1^X\). 
\begin{cor}\label{cor: short cor}
	The random variables \(\{\tau_1^{(p)}\}_{p\geq 1}\) converge in distribution to \(\tau_1^X\). 
\end{cor}
\begin{proof}
	By Corollary~\ref{cor: aln222} the probabilities 
	\begin{align}
		&\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell(\ell_0,j),j),\tau_1^{(p)}\leq E^\lambda
            	 \mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
	 	%
		\\&\to  
			\mathbb P(\bs X(\tau_1^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j),\tau_1^{(p)}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\nonumber
	\end{align}
	By the law of total probability and the convergence above, 
	\begin{align}
		&\mathbb P( \tau_1^{(p)}\leq E^\lambda
		\mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
		\\&=\sum_{\ell\in\{\ell_0-1,\ell_0,\ell_0+1\}\cap\mathcal K}\sum_{j\in\calS}\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell,j),\tau_1^{(p)}\leq E^\lambda
		\mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
	 	%
		\\&\to  
		\sum_{\ell\in\{\ell_0,\ell_0+1\}\cap\{0,1,\dots,K+1\}}\sum_{j\in\calS}\mathbb P(\bs X(\tau_1^X) = (y_{\ell}, j),\tau_1^{(p)}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\nonumber
		\\&=\mathbb P(\tau^X\leq E^\lambda \mid \bs X(0) = (x_0,i) ).
	\end{align}
	Thus, the Laplace transform of \(\tau_1^{(p)}\) converges to the Laplace transform of \(\tau_1^X\). By the Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957}, then \(\{\tau_1^{(p)}\}\) converges in distribution to \(\tau_1^X\). 
\end{proof}

\section{Convergence of the QBD-RAP scheme}\label{sec: local to global}
This section is dedicated to proving a global convergence results of the QBD-RAP approximation scheme to the fluid queue. Ultimately, we prove the weak convergence (in space and time) of the QBD-RAP approximation scheme to the distribution of the fluid queue. The structure of the argument is to first show a convergence result for the Laplace transform with respect to time of the distributions of the QBD-RAP and fluid queue at the \(n\)th orbit restart epoch. We then prove a convergence result of the QBD-RAP and fluid queue between the \(n\)th and \(n+1\)th orbit restart epoch. Summing over the number of orbit restart epochs, \(n\), and via the Dominated Convergence Theorem, we claim a convergence of the Laplace transforms with respect to time of the QBD-RAP approximation and fluid queue. Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to claim a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.

\subsection{At the \(n\)th orbit restart epoch}\label{sec: nth change}
% So far we have shown error bounds for the QBD-RAP approximation on the event that the fluid queue and QBD-RAP remain in the same band/level (%Corollary~\ref{cor: lst diff} and 
% Theorem~\ref{thm: a thm!}) and also immediately upon exiting the band (Corollary~\ref{cor: aln222}). In the next three subsections we extend this to global convergence of the approximation. To do so, we partition the approximation and the fluid queue by the number of level changes. The local convergence on a given level, which we just proved, is then used to claim that each term in the partition converges. What remains is to argue that we can swap a limit and countable sum which we do by the Dominated Convergence Theorem.

% Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
% \[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
% be the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary. To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed. Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi^{(p)}(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the value of the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\). The subscript \(\bs \alpha^{(p)}\) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\). The process \(\{\Ydp(n)\}_{n\geq 0}\) is a discrete-time Markov chain, which is time-homogeneous for \(n\geq 1\).

% % The process \(\{L^{(p)}(\tau_n^{(p)})\}_{n\geq 0}\) is the level process of the (continuous-time) QBD-RAP process observed at changes of level. When \(\tau_n^{(p)}\) is a change of level, or the time when the process leaves a boundary, the value of orbit process at these times is \(\bs A^{(p)}(\tau_n^{(p)}) = \bs \alpha^{(p)}\). At time \(\tau_0^{(p)} =0\), \(\bs A^{(p)}(\tau_0^{(p)})=\bs A^{(p)}(0) = \bs   a_{\ell_0,i}^{(p)}(x_0)\). The process \(\{\Ydp(t)\} = \{(L^{(p)}(\tau_n^{(p)}),\phi^{(p)}(\tau_n^{(p)}))\}_{n\geq 0}\) is a time-inhomogeneous discrete-time QBD. Further, the process \(\{\phi^{(p)}(\tau_n^{(p)})\}_{n\geq 0}\) is a time-inhomogeneous discrete-time Markov chain on the state space \(\mathcal S\). For \(\bs   a_{\ell_0,i}^{(p)}(x_0) = \bs \alpha^{(p)}\), or for index \(n\geq 1\), then \(\{L^{(p)}(\tau_n^{(p)}),\phi^{(p)}(\tau_n^{(p)})\}_{n\geq 1}\), and \(\{\phi^{(p)}(\tau_n^{(p)})\}_{n\geq 1}\) are time-homogenous.

% Let \(\{\tau_n^X\}_{n\geq 0}\), be the sequence of (stopping) times with \(\tau_0^X=0\), and 
% \[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\}.\]
% For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or the process leaves a boundary, for the \(n\)th time.

For \(n\geq 1\), consider the Laplace transform 
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\in \wrt t
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \wrt t \nonumber 
	 %
	 \\&=\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^{\lambda}
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\label{eqn: aaaaaaccccc} 
	%
\end{align}
% \begin{align}
% 	&\int_{t=0}^\infty e^{-\lambda t} \mathbb P(L^{(p)}(\tau_n^{(p)}) = \ell, \phi^{(p)}(\tau_n^{(p)}) = j_n, \tau_{n}^{(p)}\in \wrt t
% 	 \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
% 	 \\&\qquad{} \phi^{(p)}(0)=i) \wrt t \nonumber 
% 	 %
% 	 \\&=\mathbb P(L^{(p)}(\tau_n^{(p)}) = \ell, \phi^{(p)}(\tau_n^{(p)}) = j_n, \tau_{n}^{(p)}\leq E^{\lambda}
% 	 \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
% 	 \\&\qquad{} \phi^{(p)}(0)=i),\nonumber 
% 	%
% \end{align}
which is the Laplace transform of the time until the \(n\)th orbit restart epoch of the QBD-RAP on the event that the level and phase at the \(n\)th orbit restart epoch are \(\ell\) and \(j_n\), respectively, given that the initial level and phases are \(\ell_0\) and \(i\), respectively, and the initial orbit is \(\bs a_{\ell_0,i}^{(p)}(x_0)\). Partitioning on the time of the first orbit restart epoch, \(\tau_1\), and the level and phase at this time, then (\ref{eqn: aaaaaaccccc}) is equal to 
\begin{align}
	&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda ) \nonumber
	 %
	 \\&\times \mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
	 \mid \bs Y(0)=\bs y_0^{(p)}).
	%  \\&\times \mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
	%  \\&\qquad{} \phi^{(p)}(0)=i).
	 \label{eqn: mnebrb2}
\end{align}
An application of Corollary \ref{cor: aln222} to the expression on the second line of (\ref{eqn: mnebrb2}) states, 
\begin{align}
	&\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \nonumber
	%  &\lim_{p\to\infty}\mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
	%  \\&\qquad{} \phi^{(p)}(0)=i) \nonumber
	 %
% 	 	\\&\to \begin{cases}
% 			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j_1), \tau_{1}^X\leq E^\lambda 
%             	 \mid \bs X(0) = (x_0, i)) 
% 	 %
% 	 &\text{ for }j_1\in\calS_+,\, \ell_1=\ell_0+1\\ 
% 	%  &\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
% 	%  \mid \bs Y^{(p)}(0)=(\ell_0, \bs  a_{\ell_0,i}^{(p)}(x_0), i)) \nonumber
% 	%  &\lim_{p\to\infty}\mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
% 	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
% 	%  \\&\qquad{} \phi^{(p)}(0)=i) \nonumber
% 	 %
% 	 	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0}, j_1), \tau_{1}^X\leq E^\lambda 
%             	 \mid \bs X(0) = (x_0, i)) & \text{for }j_1\in\calS_-,\, \ell_1=\ell_0-1. 
% \end{cases},\nonumber 
	\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell+1(j_1\in\calS_-)}, j_1), \tau_{1}^X\leq E^\lambda 
	\mid \bs X(0) = (x_0, i)) 
\end{align}
for \(i\in\calS\), \(j_1\in\calS_+\cup\calS_-\), \(\ell_0\in\mathcal K\), \(x_0\in\calD_{\ell_0}\).

We now turn our attention to the first factor in the summands of (\ref{eqn: mnebrb2}). For a given \(j_1\in\calS_+\cup\calS_-\) and \(\ell_1\in\mathcal K\),  consider
\begin{align}
	&\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda) \nonumber 
	 % 
	\\&=\mathbb P(\Ydp(n-1) = (\ell, j_n), \tau_{n-1}^{(p)}\leq E^\lambda 
	 \mid \bs Y(0)=(\ell_1,\bs \alpha, j_1))\label{eqn: dskvnaSF}
\end{align}
by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution.

Let 
\begin{align}\label{eqn: paths set1}
	\mathcal P^n_{\ell_0,\ell_n}&=\left\{(\ell_1,\dots,\ell_{n-1}) \in \mathcal K^{n-1}\Bigg| \begin{array}{c}  \ell_{r-1}=\ell_r=0 \mbox{ or } \ell_{r-1}=\ell_r=K
		\\ \mbox{ or }|\ell_{r-1}-\ell_r|=1,r = 1,\dots,n\end{array}\right\}.
\end{align}
The set \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{L(\tau_m)\}_{m=2}^{n-1}\) may take on a sample path which starts in level \(\ell_0\), ends in level \(\ell\) and has \(n\) orbit restart epochs. %Similarly, \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{X(\tau_m^X)\}_{m=2}^{n-1}\) may take on a sample path which starts in band \(\ell_0\), ends in band \(\ell\) at time \(\tau_n^X\).

Thus, by partitioning on the phases and the levels at the times \(\tau_m\), \(m=2,\dots,n-1\),and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is  
	\begin{align}
		\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}&\mathbb P(\Ydp(m) =(\ell_m, j_m), 
		\tau_{m}^{(p)}\leq E^\lambda 
		\mid  \nonumber 
\Ydp (m-1) = (\ell_{m-1}, j_{m-1}), \\&\tau_{m-1}^{(p)}\leq E^\lambda ) \nonumber
		%
		\\=\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}&\mathbb P(\Ydp(1) =(\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
            	 \mid  
	 	\bs Y^{(p)}(0) = (\ell_{m-1},\bs \alpha, j_{m-1})),  \label{eqn: 161222}
\end{align}
by the time-homogeneity property of the QBD-RAP and the memoryless property of \(E^\lambda\) and where we define \(\ell_n=\ell\) for notational convenience. %Each factor in the product is 
% \begin{align}
% 	&\mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda 
%             	 \mid \Ydp(m-1) = (\ell_{m-1}, \nonumber
% 	 	 j_{m-1}), \tau_{m-1}^{(p)}\leq E^\lambda) 
% 	\\&=\mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda 
%             	 \mid \Ydp(0) = (\ell_{m-1}, j_{m-1}), \tau_{m-1^{(p)}}=0)\nonumber
% 	\\&=\mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
%             	 \mid \Ydp(0) = (\ell_{m-1}, j_{m-1}))\label{eqn: kk}
% \end{align}
% by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution. 
We can apply Corollary \ref{cor: aln222} to the factors of the summands in (\ref{eqn: 161222}) and conclude 
\begin{align}
	& \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
            	 \mid \bs Y^{(p)}(0) = (\ell_{m-1},\bs \alpha, j_{m-1}))\nonumber
	%
	\\&\to \mathbb P(\bs X(\tau_1^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{1}^X\leq E^\lambda 
	\mid \bs X(0) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1})).  \label{eqn:kknvf fawoprgj v2}
\end{align}

By the time-homogeneous property of the fluid queue and the memoryless property of the exponential distribution, (\ref{eqn:kknvf fawoprgj v2}) is equal to 
\begin{align}
	&\mathbb P(\bs X(\tau_m^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{m}^X\leq E^\lambda 
		\mid \bs X(\tau_{m-1}) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1}), \nonumber
	\\&\tau_{m-1}^X\leq E^\lambda ).\label{eqn: lim 12}
\end{align}
% \begin{align}
% 	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{1}^X\leq E^\lambda 
%             	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
% 	%  \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
%     %         	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}))\nonumber 
% 	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
%             	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \nonumber 
% 	\tau_{m-1}^X\leq E^{\lambda}).
% \end{align}
% Similarly, 
% \begin{align}
% 	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
%             	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
% 	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}}, j_m), \tau_{m}^X\leq E^\lambda 
%             	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda).
% \end{align}

Thus, returning to (\ref{eqn: 161222}) and taking the limit as \(p\to\infty\),  
\begin{align}
	&\lim_{p\to\infty}\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}\mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda \nonumber
            	 \mid \bs Y^{(p)}(0) = (\ell_{m-1},\bs\alpha, j_{m-1})) \nonumber 
	%
	\\ &=\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}\lim_{p\to\infty} \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
            	 \mid \bs Y^{(p)}(0) = (\ell_{m-1}, \bs\alpha,
	 	 j_{m-1}) ),
	 \label{eqn: 161222b}
\end{align}
where we may swap the limit and the sums as they are finite, and we can swap the limit and the product since all the limits exist and the product is finite. Substituting the limits (\ref{eqn: lim 12}) into (\ref{eqn: 161222b}) gives 
\begin{align}
	&\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}} \prod_{m=2}^{n}\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m}+1(j_{m}\in\mathcal S_-)}, j_m), \tau_{m}^X\leq E^\lambda\mid \nonumber
            	\\&\qquad\qquad  \bs X(\tau_{m-1}^X)=(y_{\ell_{m-1}+1(j_{m-1}\in\mathcal S_-)},j_{m-1}),\tau_{m-1}^X\leq E^\lambda),\nonumber
		%
		\\&= \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
		j_n), \tau_{n}^X\leq E^\lambda \mid \bs X(0)=(y_{\ell_{1}+1(j_{1}\in\mathcal S_-)},
		j_{1}), 
		\tau_1^X\leq E^\lambda),
\end{align}
by the strong Markov property of the fluid queue and the Law of total probability.

Returning now to (\ref{eqn: mnebrb2}) and taking the limit as \(p\to \infty \),  
\begin{align}
	&\lim_{p\to\infty}\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \Ydp(1)=(\ell_1,j_1), \tau_{1}^{(p)}\leq E^{\lambda}) \nonumber 
	 %
	 \\& \quad\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \nonumber
	 %
	 \\&= \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K} \lim_{p\to\infty} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda) \nonumber
	 %
	 \\& \quad\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)})\nonumber
	 %%
	 %%
	 \\&= \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K} \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
		j_n), \tau_{n}^X\leq E^\lambda \mid \nonumber
		\bs X(\tau_1^X)=(y_{\ell_{1}+1(j_{1}\in\mathcal S_-)},
		j_{1}),
		\\&\qquad{}\tau_1^X\leq E^\lambda) \nonumber
	 %
	 \mathbb P(\bs X(\tau_1^X)=(y_{\ell_{1}+1(j_{1}\in\mathcal S_-)},
		j_{1}),\tau_1^X\leq E^\lambda
		\mid \bs X(0)=(x_0, i)) \nonumber
	%
	\\&=\mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
		j_n), \tau_{n}^X\leq E^\lambda
		\mid \bs X(0)=(x_0, i))
	 \label{eqn: mnebrb22}
\end{align}
where the swapping of limits and sums in the first equality is justified as the sums are finite, the swapping limits and products in the first equality is justified as the product is finite and all limits exist, and the last inequality is the Law of total probability.

Hence, we have proved the following result 
\begin{lem}\label{lem: kKKJJJF}
	For all \(\ell,\ell_0\in\mathcal K,\) \(i,j_n\in\mathcal S\), \(x_0\in\mathcal D_{\ell_0,i}\), \(n\geq 1\), then, as \(p\to\infty\),
	\begin{align}
		&\mathbb P(\Ydp(n) = (\ell,j_n), \tau_{n}^{(p)}\leq E^{\lambda}
		\mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\nonumber 
		%
		\\&\to \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
		j_n), \tau_{n}^X\leq E^\lambda
		\mid \bs X(0)=(x_0, i)).
	\end{align}
\end{lem} 
Lemma~\ref{lem: kKKJJJF} is what is required to proceed with the proof of convergence of the QBD-RAP scheme. However, at this point we also point out the following corollary, which may be of interest in other contexts. 
\begin{cor}
	The random variables \(\{\tau_n^{(p)}\}_{p\geq 1}\) converge in distribution to \(\tau_n^X\). 
\end{cor}
\begin{proof}
	The proof follows the same arguments as the proof of Corollary~\ref{cor: short cor}. In Lemma~\ref{lem: kKKJJJF} sum over \(\ell\) and \(j_n\), then apply the Extended Continuity Theorem \cite[Chapter XIII, Theorem 2a]{feller1957}.
\end{proof}

\subsection{Between the \(n\)th and \(n+1\)th orbit restart epochs}\label{sec: between n and np1}
In the last section we proved a convergence of the QBD-RAP to the fluid queue \emph{at} the \(n\)th orbit restart epoch. However, we ultimately want to make a convergence statement about the QBD-RAP scheme at any time. The next step is therefore to show a convergence of the QBD-RAP to the fluid queue between the \(n\)th and \(n+1\)th orbit restart epoch, after which we sum over \(n\) to prove a convergence result independent of the number of orbit restart epochs.

Let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\). Consider the Laplace transform 
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid 
	\bs Y^{(p)}(0)=\bs y_0^{(p)})\psi(x)\wrt t, \label{eqn: aaabbababbaa}
	% \nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_\ell}\mathbb P(L^{(p)}(t) = \ell, \widetilde X^{(p)}(t)\in\wrt x, \phi^{(p)}(t) = j, \tau_{n}^{(p)}\leq t<\tau_{n+1}^{(p)} \mid L^{(p)}(0)=\ell_0, 
	% 	\\&\qquad \bs A^{(p)}(0)=\bs a_{\ell_0,i}^{(p)}(x_0), \phi^{(p)}(0)=i)\psi(x)\wrt t.
\end{align}
where \(\psi:\mathbb R\to \mathbb R\) is a bounded function.

Partitioning (\ref{eqn: aaabbababbaa}) on the time of the \(n\)th orbit restart epoch and the phase and level at this time gives
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t} 
	\int_{x\in\calD_{\ell,j}}
	\int_{u_n=0}^t\sum_{j_n\in\mathcal S} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
	t \in(u_n, \tau_{n+1}^{(p)}] \mid\Ydp(n) = (\ell,j_n), \nonumber 
	\\&\quad \tau_{n}^{(p)}= u_n)\psi(x) \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\in \wrt u_n 
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) 
	  \wrt t \nonumber 
	 %
	 \\&= \sum_{j_n\in\mathcal S}
	\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), 
	t\in\mathcal T_{0}^{(p)} \mid\nonumber 
	  \bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t \psi(x)  
	  \\&\quad{} \mathbb P(\Ydp(n) = (\ell,j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \label{eqn: l whkrqvlkjbrd}
\end{align}
by the time homogenous property of the QBD-RAP, the memoryless property of the exponential distribution, and the convolution theorem of Laplace transforms. The swap of integrals and sums is justified by the Fubini-Tonelli Theorem. We recognise the probability 
\begin{align}
	\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \label{eqn: Kkk nj aa w}
\end{align}
as that appearing in Lemma \ref{lem: kKKJJJF}, hence (\ref{eqn: Kkk nj aa w}) converges to 
\begin{align}
	\mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
		j_n), \tau_{n}^X\leq E^\lambda
		\mid \bs X(0)=(x_0, i)) \label{eqn: ashverkjbvjhkhdsknacva}
\end{align}
as \(p\to\infty\).

Now consider the expression 
\begin{align}
	&\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j),  
	t\in\mathcal T_0^{(p)} \mid \bs Y(0) = (\ell,\bs\alpha, j_n)
	 	 )\wrt t  \psi(x)  \label{eqn: vajkl;vJK:GSKJ}
\end{align}
which appears as part of (\ref{eqn: l whkrqvlkjbrd}). We can rewrite (\ref{eqn: vajkl;vJK:GSKJ}) as  
\begin{align}
	 &\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
	t\in\mathcal T_0^{(p)}\mid \Ydp(0) = (\ell,j_n))\wrt t \psi(x) \nonumber 
	 \\&= \int_{x\in\calD_{\ell,j}} \widehat f^{\ell,(p)} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x \label{eqn: Jk jJ KK}
\end{align}
Applying Theorem~\ref{thm: a thm!}, then (\ref{eqn: Jk jJ KK}) converges to 
\begin{align}
	&\int_{x\in\calD_{\ell,j}} \widehat \mu^{\ell} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x \nonumber 
	\\&= \int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_0^X \mid \bs X(0) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n) ) \psi(x). \label{eqn: lkHAHJFHKJ J J H}
\end{align}
Since the fluid queue is time-homogeneous and \(E^\lambda\) memoryless, then (\ref{eqn: lkHAHJFHKJ J J H}) is equal to 
\begin{align}
	&\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x,j), t\in\mathcal T_n^{X} \mid  
	\bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)},j_n), \tau_n^X\leq E^\lambda ) \psi(x).  \label{eqn: lkHAHJFHKJ J J 22}
\end{align}
Therefore, we have shown (\ref{eqn: vajkl;vJK:GSKJ}) converges to (\ref{eqn: lkHAHJFHKJ J J 22}) as \(p\to\infty\).

Returning to the right-hand side of (\ref{eqn: l whkrqvlkjbrd}) and taking the limit as \(p\to\infty\), 
\begin{align}
	&\lim_{p\to\infty} \sum_{j_n\in\mathcal S}
	\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), 
	t\in\mathcal T_0^{(p)} \mid \nonumber 
	 \bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t \psi(x)
	 \\&\qquad{}  \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}).\label{eqn: eieieie}
	 \\&=\sum_{j_n\in\mathcal S}\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} \mid \bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n),\nonumber 
	 \tau_n^X\leq E^\lambda ) 
	 \\&\quad{} \psi(x)\mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
		j_n), \tau_{n}^X\leq E^\lambda
		\mid \bs X(0)=(x_0,i))\nonumber
	%
	\\&= \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} 
	\mid X(0)=x_0, \varphi(0)=i)\psi(x)\wrt t,
\end{align}
where the first equality holds since the sum is finite and the limits exist, and the second equality holds from the law of total probability.

Hence, we have shown the following result 
\begin{lem}\label{lem: LAkAKFnvnb mav h}
	For \(\ell,\ell_0\in\mathcal K\), \(i,j\in\calS\), \(n\geq 0\), then, as \(p\to\infty\), 
	\begin{align} 
		&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), t\in\mathcal T_n^{(p)} \mid  \nonumber 
		\bs Y(0) = \bs y_0)\psi(x)\wrt t \nonumber
		%
		\\&\to\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j_n}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} 
		\mid \bs X(0)=(x_0,i))\psi(x)\wrt x\wrt t.
	\end{align}
\end{lem}

\subsection{A domination condition}
% We now want to sum the expressions in Lemma~\ref{lem: LAkAKFnvnb mav h} over \(n\) to obtain a result which is independent of the number of orbit restart epochs. To do so, we use the Dominated Convergence Theorem. We prove an appropriate domination condition below.
Our aim is to prove convergence of 
\begin{align} 
	&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j) \mid  \nonumber 
	\bs Y(0) = \bs y_0)\psi(x)\wrt t \nonumber
	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), t\in\mathcal T_n^{(p)} \mid  \nonumber 
	\bs Y(0) = \bs y_0)\psi(x)\wrt t \nonumber
	\\&=\sum_{n=0}^\infty\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), t\in\mathcal T_n^{(p)} \mid   
	\bs Y(0) = \bs y_0)\psi(x)\wrt t \label{eqn:GHJKLKJHGHJKL}
\end{align}
where the swap of the intergals and sums is justified by the Fubini-Tonelli Theorem. Now, (\ref{eqn:GHJKLKJHGHJKL}) is an infinite sum of terms appearing in Lemma~\ref{lem: LAkAKFnvnb mav h}. Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obtain the desired result. To this end, we now show a domination condition in Corollary~\ref{vcor: cdks d} so that we may apply the Dominated Convergence Theorem. %The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the aforementioned swap of sum and limit.

% To extend Lemma~\ref{lem: LAkAKFnvnb mav h} to show a global convergence result we use the Dominated Convergence Theorem. Here we show some geometric bounds on the probability of \(n\) level changes of the QBD-RAP and fluid queue, which ultimately serve as a dominating function in the Dominated Convergence Theorem.

%Let \(c_{min} = \min\limits_{i\in\mathcal S_+\cup \mathcal S_-}|c_i|\).

\begin{lem}\label{lem: another bound}
	For all \(i\in\mathcal S_+\cup \calS_-\), and \(n\geq 2\),
	\begin{align}
		&\mathbb P(\tau_n\leq E^\lambda \mid \phi(\tau_{n-1})=i, \tau_{n-1}\leq  E^\lambda ) \leq b,
	\end{align}
	where 
	\[b = \min\left\{1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,\cfrac{q}{q+\lambda}\right\}\]
	and  
	\[|r_2|\leq 2G\cfrac{\var \left(Z\right)}{\varepsilon^2} + 2L\varepsilon.\]
\end{lem}
Note that \(b\) and \(r_1\) depend on \(p\) which has been suppressed to simplify notation. When explicitly needed, we use a superscript \(p\) to denote this dependence.  
\begin{proof}
	For the QBD-RAP, orbit restart epochs can only occur when \(i\in\mathcal S_+\cup\calS_-\).

	Suppose that the phase at time \(\tau_{n-1}\) is \(i\in\mathcal S_+\) and that at time \(\tau_{n-1}\) the QBD-RAP is not at a boundary. The arguments for an initial phase \(i\in\mathcal S_-\) are analogous. For an orbit restart epochs to occur, either; 
	\begin{enumerate}
		\item the QBD-RAP remains in phase \(i\) until the orbit restart epoch occurs, or
		\item the QBD-RAP changes phase before there is an orbit restart epoch, after which the orbit restart epoch occurs eventually. 
	\end{enumerate}

	Hence, for sample paths which contribute to the Laplace transform, one of two things must happen, either; 
	\begin{enumerate}
		\item the phase remains \(i\) until there is an orbit restart epoch and \(E^\lambda\) does not occur before the orbit restart epoch, or, \label{item: 1}
		\item the phase changes before there is an orbit restart epoch and \(E^\lambda\) does not occur before the orbit restart epoch. \label{item: 2}
	\end{enumerate}

	The probability of~\ref{item: 1}~is 
	\begin{align}
		\int_{x=0}^\infty \bs \alpha e^{\bs{S}x}\bs s e^{(T_{ii}-\lambda)x/|c_i|}\wrt x 
		&= e^{(T_{ii}-\lambda)\Delta/|c_i|} + r_2,\label{eqn: bnd this 12345}
	\end{align}
	by Lemma~\ref{lemma:bound}.

	The probability of~\ref{item: 2}~is 
	\begin{align}
		\nonumber\int_{x=0}^\infty \bs \alpha e^{\bs{S}x}\bs e e^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x 
		&= \int_{x=0}^{\Delta+\varepsilon} \bs \alpha e^{\bs{S}x}\bs ee^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x 
%		\\\nonumber&{}+\int_{x=\Delta}^{\Delta+\varepsilon}\bs \alpha e^{\bs{S}x}\bs ee^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x 
		\\&{}+\int_{x=\Delta+\varepsilon}^\infty \bs \alpha e^{\bs{S}x}\bs e e^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x.\label{eqn: bnd this 1234}
	\end{align}
	Now, since \(\bs \alpha e^{\bs{S}x}\bs e\leq 1\) for \(x\leq \Delta+\varepsilon\) then the first term on the right-hand side of (\ref{eqn: bnd this 1234}) is less than or equal to 
	\[\displaystyle \int_{x=0}^{\Delta+\varepsilon} e^{(T_{ii}-\lambda)/|c_i|x}(-T_{ii}/|c_i|)\wrt x \leq \int_{x=0}^{\Delta+\varepsilon} e^{T_{ii}/|c_i|x}(-T_{ii}/|c_i|)\wrt x = 1-e^{T_{ii}/|c_i|(\Delta+\varepsilon)}.\]
	By Chebyshev's inequality, \(\bs \alpha e^{\bs{S}x}\bs e\leq \cfrac{\var(Z)}{\varepsilon^2}\) for \(x> \Delta+\varepsilon\), hence the second term on the right-hand side of (\ref{eqn: bnd this 1234}) is less than or equal to 
	\[\displaystyle \int_{x=\Delta+\varepsilon}^\infty \cfrac{\var(Z)}{\varepsilon^2} e^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x \leq  \cfrac{\var(Z)}{\varepsilon^2}.\]
	Putting these together, then the right-hand side of (\ref{eqn: bnd this 1234}) is less than or equal to
	\begin{align}
		1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|} + \cfrac{\var(Z)}{\varepsilon^2}.\label{eqn: the big dog}
	\end{align}	
	Combining (\ref{eqn: bnd this 12345}) and (\ref{eqn: the big dog}), then \(\mathbb P(\tau_n\leq E^\lambda  \mid \phi(\tau_{n-1})=i , \tau_{n-1}\leq  E^\lambda)\) is less than or equal to 
	\begin{align}
		&e^{(T_{ii}-\lambda)\Delta/|c_i|} + |r_2| + 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|} + \cfrac{\var(Z)}{\varepsilon^2}\nonumber
		\\&= 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|}\left[1-e^{(-T_{ii}\varepsilon-\lambda \Delta)/|c_i|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2| \nonumber 
		\\&= 1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,
	\end{align}
	since \(-T_{ii}/|c_i|\leq q\) and \(\lambda \Delta/|c_i| \leq \lambda \Delta/c_{min}\) for all \(i\in\mathcal S_+\cup \calS_-\).

	Now consider the QBD-RAP at a boundary. To leave the boundary there must be at-least one change of phase before \(E^\lambda\). By a uniformisation argument, the probability of at-least one change of phase before \(E^\lambda\) is less than or equal to \(q/(q+\lambda)\). 
\end{proof}

\begin{lem}\label{lem: another bound 2}
	For \(n\geq 2\), \(i\in\mathcal S_+\cup \calS_-\), 
	\begin{align}
		\mathbb P(\tau_n \leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \leq b^{n-1}.
	\end{align}
\end{lem}
\begin{proof}
	The proof is by induction.

	For the base case, set \(n=2\) and apply Lemma~\ref{lem: another bound}.

	Now, assume the induction hypothesis \(\mathbb P(\tau_{n-1} \leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \leq b^{n-2}\) for arbitrary \(n\geq 3\).

	Since \(\{\tau_{n-1}\leq E^\lambda\}\) is a subset of \(\{\tau_{n}\leq E^\lambda\}\), then 
	\begin{align}
		 \mathbb P(   \tau_n \leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)
		&= \mathbb P(   \tau_n \leq E^\lambda,  \tau_{n-1} \leq E^\lambda  \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda). \label{eqn: eheh}
	\end{align}
	Now partition (\ref{eqn: eheh}) on the phase at time \(\tau_{n-1}\),
	\begin{align}
		\nonumber&\sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_n \leq E^\lambda,  \tau_{n-1} \leq E^\lambda,\phi(\tau_{n-1}) = j_{n-1} \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)
		\\&\nonumber=  \sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_{n}\leq E^\lambda\mid   \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1} \leq E^\lambda)
		\\&\qquad\times\mathbb P( \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda), \label{eqn: kekew}
	\end{align}
	by the strong Markov property of the QBD-RAP and the fact that \(\bs A(\tau_{n-1})=\bs \alpha\).

	By Lemma~\ref{lem: another bound} (\ref{eqn: kekew}) is less than or equal to 
	\begin{align}
		&\sum_{j_{n-1}\in\mathcal S} b
		\mathbb P(  \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda\mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \nonumber
		\\&= b
		\mathbb P( \tau_{n-1}\leq E^\lambda\mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)\nonumber
		\\&\leq b\cdot b^{n-2},
	\end{align}
	by the induction hypothesis, and this completes the proof. 
\end{proof}
%

\begin{cor}\label{vcor: cdks d}
	For \(\bs y_0=(\ell_0,\bs a_{\ell_0,i}(x_0),i),\, \lambda>0,\, \ell_0,\ell\in\mathcal K,\,i,j\in\mathcal S\) and any bounded function \(\psi\leq F\), then
	\begin{align}
		\Bigg|&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}(t) \in (\ell,\wrt x, j), t\in\mathcal T_n \mid 
		\bs Y(0)=\bs y_0)
		\psi(x) \wrt t \Bigg|
		\leq \cfrac{F b^{n-1}}{\lambda}.\label{eq: JjbCS  B k }
	\end{align}
\end{cor}
\begin{proof}
	First, since \(|\psi(x)|\leq F\), then the left-hand side of (\ref{eq: JjbCS  B k }) is less than or equal to 
	\begin{align}
		&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}(t) \in (\ell,\wrt x,j), t\in\mathcal T_n \mid  \nonumber 
		\bs Y(0)=\bs y_0)F\wrt t \nonumber
		\\&=\int_{t=0}^\infty e^{-\lambda t} \mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \mid \bs Y(0)=\bs y_0)F\wrt t. \label{eqn: alalal}
	\end{align}
	Partitioning on the time of the \(1\)st orbit restart epoch, \(\tau_1\), and the phase and level at time \(\tau_1\), then (\ref{eqn: alalal}) is equal to 
	\begin{align}
		&\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \nonumber
		\mid \Yd(1)=(\ell_1,j_1),  \tau_1=u_1)\nonumber
		\\&\quad\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
		\mid \bs Y(0)=\bs y_0)F  \wrt t\nonumber
		%
		\\&=\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t-u_1\in\mathcal T_{n-1} \nonumber
		\mid 
		\bs Y(0)=(\ell_1,\bs \alpha,j_1))
		\\&\quad\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
		\mid \bs Y(0)=\bs y_0)F  \wrt t, \label{eqn fLks}
	\end{align}
	by the time-homogeneous property of the QBD-RAP. By the convolution theorem for Laplace transforms, (\ref{eqn fLks}) is equal to 
	\begin{align}
		&\sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
		\mid \bs Y(0)=(\ell_1,\bs \alpha, j_1)) \wrt t 
		\\&\quad{}\int_{u_1=0}^\infty e^{-\lambda u_1}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
		\mid \bs Y(0)=\bs y_0)F  \nonumber 
		%
		%
		\\&=\sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t 
		\\&\quad{}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\leq E^\lambda
		\mid \bs Y(0)=\bs y_0)F. \label{eqn fLksoioi}
	\end{align}

	The expression 
	\begin{align}
		&\int_{t=0}^\infty e^{-\lambda t} \mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n\nonumber
		\mid \bs Y(0)=(\ell_1,\bs \alpha,j_1)) \wrt t
		%
		\\&\leq \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\tau_{n}\leq t \nonumber
		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t
		%
		%
		\\&\leq b^{n-1}\int_{t=0}^\infty e^{-\lambda t} \wrt t \nonumber 
		\\&=b^{n-1}\cfrac{1}{\lambda}, \label{eq: bndhs vja sdh}
	\end{align}
	by Lemma~\ref{lem: another bound 2}.

	Using the bound (\ref{eq: bndhs vja sdh}) in (\ref{eqn fLksoioi}) gives 
	\begin{align}
		&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K} b^{n-1}\cfrac{1}{\lambda} 
		\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\leq E^\lambda
		\mid \bs Y(0)=\bs y_0)F\leq b^{n-1}\cfrac{1}{\lambda}F, \label{eqn fLksoioi2}
	\end{align}
	by the law of total probability. This concludes the proof. 
\end{proof}

\subsection{Global convergence}\label{sec: sub global}
Finally, we combine the convergence result of Lemma~\ref{lem: LAkAKFnvnb mav h} and the domination condition from Corollary~\ref{vcor: cdks d} via the Dominated Convergence Theorem to claim convergence of the Laplace transform of the QBD-RAP given by 
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y}^{(p)}(t) = (\ell,\wrt x,j)\mid \bs Y^{(p)}(0)=\bs y_0^{(p)})  \psi(x)\wrt t. \label{eqn: LLLLaaakkkss}
\end{align}
Partitioning (\ref{eqn: LLLLaaakkkss}) on the number of orbit restart epochs by time \(t\) gives
\begin{align}
	& \int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\sum_{n=0}^\infty \mathbb P(  {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid\nonumber 
	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \psi(x) \nonumber
	\\&=\sum_{n=0}^\infty\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}} \mathbb P( {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid  
	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \psi(x). \label{eqn: KLDNVnav}
\end{align}
We can justify the swap of the sum and integrals since 
\begin{align}
	& \int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\sum_{n=0}^\infty \mathbb P(  {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid\nonumber 
	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t |\psi(x)| \nonumber
	\\&\leq \int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\sum_{n=0}^\infty \mathbb P(  {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid\nonumber 
	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t F \nonumber
	\\&\leq \int_{t=0}^\infty e^{-\lambda t}\wrt t F \nonumber
	\\&\leq \cfrac{1}{\lambda} F<\infty \nonumber
\end{align}
so the Fubini-Tonelli Theorem applies.

By Lemma (\ref{lem: LAkAKFnvnb mav h}), each term in the sum (\ref{eqn: KLDNVnav}) converges. Furthermore, for \(n\geq 1\), each term is dominated by \(\left(b^{(p)}\right)^{n-1}F/\lambda\), from Corollary~\ref{vcor: cdks d}. The dominating terms \(\left(b^{(p)}\right)^{n-1}F/\lambda\) depend on \(p\) and may not be summable. However, for \(p\) sufficiently large, there exists a \(p_0<\infty\) and a \(B\) with \(B<1\) such that \(b^{(p)}<B\) for all \(p>p_0\). Hence we can apply the Dominated Convergence Theorem to (\ref{eqn: KLDNVnav}) and claim that  
\begin{align}
	&\lim_{p\to\infty} \sum_{n=0}^\infty\int_{x\in\mathcal D_{\ell,j}}\int_{t=0}^\infty e^{-\lambda t} \mathbb P( {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid \nonumber 
	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \psi(x)  \nonumber
	%
	% \\&= \sum_{n=0}^\infty\int_{x\in\mathcal D_{\ell,j}}\lim_{p\to\infty} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\widetilde{\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid \nonumber 
	% \bs Y^{(p)}(0)=(\ell_0,  \bs  a_{\ell_0,i}^{(p)}(x_0), i))\wrt t \psi(x) \nonumber
	%
	\\&= \sum_{n=0}^\infty\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
	\mid \bs X(0)=(x_0,i)) 
	\psi(x)\wrt t. \label{eqn:Alakskjs}
\end{align}
where we have used Lemma~\ref{lem: LAkAKFnvnb mav h}. Swapping the sum and integrals and by the law of total probability, then (\ref{eqn:Alakskjs}) is equal to 
\begin{align}
	& \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
	\mid \bs X(0)=(x_0, i))\psi(x)\wrt t \nonumber
	%
	\\&= \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}} \mathbb P(\bs X(t)\in(\wrt x, j)  
	\mid \bs X(0)=(x_0, i))\psi(x)\wrt t. \nonumber
\end{align}
The swap of the sum and integrals is justfied as 
\begin{align}
	& \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
	\mid \bs X(0)=(x_0, i))|\psi(x)|\wrt t \nonumber
	%
	\\&\leq\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
	\mid \bs X(0)=(x_0, i))\wrt t  F\nonumber
	\\&\leq\int_{t=0}^\infty e^{-\lambda t} \wrt t  F\nonumber
	\\&=\cfrac{1}{\lambda}F<\infty
\end{align}
so the Fubini-Tonelli Theorem applies.

Thus, we have shown the following result.
\begin{lem}\label{lem: KajPOw}
	For all \(\ell_0,\ell\in\mathcal K\), \(i,j\in\mathcal S\), \(x_0\in\mathcal D_{\ell_0,i}\), as \(p\to\infty\), 
	\begin{align}
		&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y}^{(p)}(t) = (\ell,\wrt x,j) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)})  \psi(x)\wrt t \nonumber
		\\&\to \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}} \mathbb P(\bs X(t)\in(\wrt x, j)  
		\mid \bs X(0)=(x_0, i))\psi(x)\wrt t. \nonumber
	\end{align}
\end{lem}

Lemma~\ref{lem: KajPOw} establishes a convergence for a given interval \(\mathcal D_\ell,\, \ell\in\mathcal K\), and phase \(j\in\mathcal S\). We now formally extend this to a global result. To do so, we find it convenient to re-write the problem in terms of expectations.

Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\), then
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \nonumber
	\bs Y (0)=\bs y_0 \Bigg] \wrt t 
	% \\&=\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\int_{x\in\calD_{L(t),\varphi(t)}}\bs A(t)\bs v_{L(t),\varphi(t)}(x)\psi(x,\varphi(t)),\varphi(t) )\wrt x  \mid \nonumber
	% \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t 
	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{\bs a\in\mathcal A}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{x\in\calD_{\ell,j}}\mathbb P( {\bs Y}(t) \in (\ell,\wrt \bs a,j) \mid \bs Y(0)=\bs y_0)\bs a\bs v_{\ell,j}(x)  \psi(x)\wrt x\wrt t \nonumber 
	\\&=\int_{t=0}^\infty e^{-\lambda t}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y} (t) \in (\ell,\wrt x,j) \mid \bs Y (0)=\bs y_0 )  \psi(x)\wrt t,
\end{align}
and the terms in the last line are those in Lemma~\ref{lem: KajPOw}.

\begin{cor}\label{cor: lk}
	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary, bounded function with \(|\psi(\cdot)|\leq F\). %Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\varphi(t)))\). 
	For each \(i\in\mathcal S\), \(\ell_0\in\mathcal K\), \(x_0\in\mathcal D_{\ell_0,i}\),
	\begin{align}
		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )  \mid \nonumber
		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber
		\\&\to \int_{t=0}^\infty e^{-\lambda t}  \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \wrt t.\nonumber
	\end{align}
\end{cor}

\begin{proof}
	Consider the left-hand side 
	\begin{align}
		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid 
		  \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
		\\&= \int_{t=0}^\infty e^{-\lambda t}\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )  1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
		\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[ \psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid 
		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t, \label{eqnL AKJF cjk ajhJK}
	\end{align}
	where the swap of the summations and integrals is on the last line justified since \(\psi\) is bounded and by the Fubini-Tonelli Theorem. By Lemma~\ref{lem: KajPOw}, for each \(\ell\in\mathcal K\), \(j\in\calS\), the terms 
	 \begin{align}
	 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j), j )\wrt x 1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t
	\end{align}
		converge to 
	\[\int_{t=0}^\infty e^{-\lambda t}  \mathbb E\left[\psi({\bs X}(t) )1(\bs X(t)\in(\mathcal D_{\ell_j},j))\mid \bs X(0)=(x_0,i) \right] \wrt t.\]

	If \(\mathcal K\) is finite, we are done upon taking the limit of (\ref{eqnL AKJF cjk ajhJK}) as \(p\to\infty\) and swapping the limit and the sums.

	If \(\mathcal K\) is countably infinite, then for a given \(k\in\mathcal K\), since \(\psi\) is bounded,
	\begin{align}
		&\Bigg|\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),\varphi(t) )  1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t\Bigg| \nonumber 
		%
		\\&\leq F \sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
		\bs Y^{(p)}(0)=\bs y_0^{(p)}\Bigg] \wrt t \nonumber 
		%
		\\&\leq F  \int_{t=0}^\infty e^{-\lambda t}\mathbb P( {\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j) \mid   
		\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \nonumber 
		%
		\\&\leq F \mathbb P(\tau_{|\ell-\ell_0|}^{(p)}\leq E^\lambda \mid
			\bs Y^{(p)}(0)=\bs y_0^{(p)}), \label{eqn:SKAM. fqj}
	\end{align}
	since, to be in level \(\ell\) after starting in level \(\ell_0\), there must be at least \(|\ell_0-\ell|\) orbit restart epochs. By Lemma~\ref{lem: another bound 2} then (\ref{eqn:SKAM. fqj}) is bounded by \(\left(b^{(p)}\right)^{|\ell-\ell_0|-1}\) for \(|\ell-\ell_0|\geq 2\) and by \(1\) otherwise. Now, choose \(p_0\) sufficiently large so that \(b^{(p)}<B<1\) for all \(p>p_0\). Therefore, for all \(p>p_0\), the terms in (\ref{eqnL AKJF cjk ajhJK}) are dominated by \(F\min\{B^{|\ell-\ell_0|-1},1\}\). Moreover, 
	\begin{align*}
		F\sum_{\ell\in\mathcal K} \min\{B^{|\ell-\ell_0|-1},1\} 
		&\leq2\sum_{n=1}^\infty B^{n-1}+1
		\\&=\cfrac{2}{1-B}+1
		\\&<\infty,
	\end{align*}
	hence the dominating terms are summable. Hence, we may apply the Dominated Convergence Theorem to swap the necessary limits and sums, from which the result follows.
\end{proof}

The Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} can now be used to claim that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.

\begin{thm}
	For all \(\bs y_0^{(p)}=(\ell,\bs a_{\ell_0,i}^{(p)}(x_0),i)\), \(\ell_0\in\mathcal K\), \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\calS\), and any bounded function \(\psi:\mathbb R\times \mathcal S \to \mathbb R\), then
	\begin{align}
			&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] \nonumber 
			\to \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \nonumber
	\end{align}
	weakly in \(t\) as \(p\to \infty\). 
\end{thm}
\begin{proof}
	Combine the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} with the convergence of Laplace transforms in Corollary~\ref{cor: lk}. 
\end{proof}

Ideally, we would like to obtain a point-wise convergence result in the variable \(t\). However, to date, this has eluded me. Although 
\[\mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0,i) \right]\] 
is a continuous function of \(t\) (it is a Feller semigroup), and, for \(p<\infty\), 
\[\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
 is continuous in \(t\), the limit 
 \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\). Nonetheless, we can claim that 
\begin{align}
		&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )    \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \nonumber
		\to \mathbb E\left[\psi({\bs X}(t) )\mid \bs X(0)=(x_0, i) \right] \nonumber
\end{align}
for almost all \(t\geq 0\). At such values of \(t\), since \(\psi\) is arbitrary and bounded, then the Portmanteau Theorem \citep[Theorem 2.1]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.

A sufficient condition to upgrade the convergence from weak to point-wise (in the variable \(t\)) is to show that for \(t\geq 0\) 
\[\sup_{p}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )  \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \leq M(t)<\infty\]
and the sequence \(\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\). That is, for every \(\varepsilon>0\) there exists a \(\delta(t,\varepsilon)>0\) and an \(p_0(t,\varepsilon)\) such that \(|t-u|<\delta(t,\varepsilon)\) implies that 
\begin{align}
	&\nonumber \Bigg|\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
	\\&{}-\mathbb E\left[\psi( \mathcal R(L^{(p)}(u),\bs A^{(p)}(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| \nonumber 
	<\varepsilon
\end{align}
for all \(p\geq p_0(t,\varepsilon)\).

% \begin{lem}
% 	\begin{align}
% 		&\Bigg|\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
% 		\\&{}-\mathbb E\left[\psi( \mathcal R(L(u),\bs A(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| < Lc_{max}|t-u| \nonumber
% 	\end{align}
% 	i.e.~uniformly Lipschitz in \(t\), and this does not depend on \(p\). 
% \end{lem}
% \begin{proof}
% 	Suppose \(i\in\calS_-\). 
% 	\begin{align*}
% 		&\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\varphi(t+u)),\varphi(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
% 		%
% 		\\&=\mathbb E\left[\int_{x\in\calD_{L(t+u),\varphi(t+u)}}\bs A(t+u)\bs v_{L(t+u),\varphi(t+u)}\psi(x,\varphi(t+u) )\wrt x   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
% 		%
% 		\\&=\int_{L(t+u),\varphi(t+u)}\bs a e^{(c_i\bs S + T_{ii}\bs I)u}\bs v_{\ell,i}(x)\psi(x,i)\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}\bs v_{\ell+1,i}(x)\psi(x,i)\wrt x 
% 		%
% 		\\&\quad{}+o(u^2)
% 		%
% 		\\&=\int_{x\in[0,\infty)} \bs a e^{(c_i\bs S + T_{ii}\bs I)u}e^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt x 
% 		%
% 		\\&\quad{}+o(u^2).
% 	\end{align*}
% 	Now, 
% 	\begin{align*}
% 		&\int_{x\in[0,\infty)} \bs a e^{(c_i\bs S + T_{ii}\bs I)u}e^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x
% 		%
% 		\\&=\int_{x\in[0,\infty)} \bs a e^{c_i\bs S u} e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x),i)\wrt x
% 		%
% 		\\&=\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
% 	\end{align*}
% 	and 
% 	\begin{align*}
% 		&\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt v\wrt x 
% 		%
% 		\\&=\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s \int_{x\in[0,\infty)} \bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt x \wrt v
% 		%
% 		\\&=\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s \int_{x\in[c_i(u-v),\infty)} \bs \alpha e^{\bs S x}\bs se^{T_{ii}(u-v)}\psi((\ell-1)\Delta+m(x),i)\wrt x \wrt v
% 		%
% 		\\&=\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s e^{T_{ii}(u-v)} \left(\psi((\ell-1)\Delta+m(\Delta-c_i(u-v)),i) + r_2\right) \wrt v
% 		%
% 		\\&=\int_{v=0}^u\bs a e^{c_i\bs Sv}c_i\bs s e^{T_{ii}u} \left(\psi((\ell-1)\Delta+m(\Delta-c_i(u-v)),i) + r_2 \right) \wrt v
% 		%
% 		\\&=\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \left( \psi((\ell-1)\Delta+m(\Delta+x-c_iu)),i) + r_2 \right) \wrt x.
% 	\end{align*}
% 	Also, for \(v\leq c_iu\leq \Delta \), 
% 	\begin{align*}
% 		\psi((\ell-1)\Delta+m(\Delta+x-c_iu)) &= \psi((\ell-1)\Delta+\Delta - |mod((\Delta+x-c_iu),2\Delta) - \Delta|)
% 		% 
% 		\\&=\psi(\ell\Delta - |x-c_iu|).
% 	\end{align*}
% 	Now, 
% 	\begin{align*}
% 		&\left|\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x \right|
% 		%
% 		\\&\leq \left|\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\wrt v\wrt x \right|F
% 		%
% 		\\&\leq \left|\sum_{j\in\calS_-} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)} \bs e\wrt v \right|F
% 		%
% 		\\&\leq \left|\sum_{j\in\calS_-} \int_{v=0}^uT_{ij}\wrt v\right|F,
% 		% 
% 		\\&= \left|\sum_{j\in\calS_-} uT_{ij} \right|F,
% 	\end{align*}
% 	similarly, 
% 	\begin{align*}
% 		&\left|\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x \right|
% 		%
% 		\\&\leq\left|\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\wrt v\wrt x \right|F
% 		%
% 		\\&\leq\left|\sum_{j\in\calS_+}\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs e\wrt v \right|F
% 		%
% 		\\&\leq\left|\sum_{j\in\calS_+}\int_{v=0}^uT_{ij}\wrt v \right|F
% 		%
% 		\\&=\left|\sum_{j\in\calS_+}uT_{ij}\right|F
% 	\end{align*}

% 	Also note
% 	\begin{align*}
% 		\int_{x\in\calD_{\ell,i}}\bs a\bs v_{\ell,i}(x)\psi(x,i)\wrt x = \int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x.
% 	\end{align*}

% 	Therefore, consider the derivative
% 	\begin{align*}
% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\varphi(t+u)),\varphi(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
% 		\\&{}-\int_{x\in\calD_{\ell,i}}\bs a\bs v_{\ell,i}(x)\psi(x,i)\wrt x \Bigg)
% 		\\&= \lim_{u\to 0^+} u^{-1}\Bigg(\int_{x\in[0,\infty)} \bs a e^{(c_i\bs S + T_{ii}\bs I)u}e^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt x 
% 		%
% 		\\&\quad{}+o(u^2)- \int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg).
% 		%
% 		\\&= \lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi((\ell-1)\Delta+m(\Delta+x-c_iu)),i)\wrt x
% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
% 		%
% 		\\&\quad{}+\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} r_2  \wrt x
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x 
% 		%
% 		\\&\quad{}+o(u^2)\Bigg).
% 	\end{align*}
% 	The second limit is less than or equal to 
% 	\begin{align*}
% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s\wrt xe^{T_{ii}u}r_2 + \left|\sum_{j\in\calS_+}uT_{ij}\right|F + \left|\sum_{j\in\calS_i}uT_{ij}\right|F + o(u^2)\Bigg) = \bs a \bs sr_2 + |T_{jj}|F.
% 	\end{align*}
% 	The first limit is 
% 	\begin{align*}
% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi((\ell-1)\Delta+m(\Delta+x-c_iu)),i)\wrt x
% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
% 		%
% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi(\ell\Delta - |x-c_iu|)\wrt x
% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
% 		%
% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\Delta)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i)\wrt x
% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi(\ell\Delta - |x-c_iu|)\wrt x
% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
% 		%
% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[0,\Delta)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i)\wrt x
% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
% 		%
% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[0,\Delta)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i)\wrt x-\int_{x=0}^\Delta\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+x,i)\wrt x
% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x -\int_{x=\Delta}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
% 		%
% 		\\&= \lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^\Delta\bs ae^{\bs Sx}\bs s\left|e^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i) - \psi(\ell\Delta+x,i)\wrt x\right|
% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\left|\psi(\ell\Delta+m(x-c_iu),i) -\psi(\ell\Delta+m(x),i)\right|\wrt x\Bigg)
% 		%
% 		\\&\leq \lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^\Delta\bs ae^{\bs Sx}\bs s L^{\psi}|T_{ii}||c_iu|
% 		+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}L^{\psi}|T_{ii}||c_iu|\wrt x\Bigg)
% 		%
% 		\\&= \lim_{u\to 0^+}u^{-1} L^{\psi}|T_{ii}||c_iu|
% 	\end{align*}
% \end{proof}

% \begin{rem}
% 	While the convergence result implies a loose bound on the rate of convergence, it applies to a wide class of QBD-RAP constructions. The main cause of the loose bound on the convergence rate is due to our reliance on Chebyshev's inequality. This is necessary as we try to make minimal assumptions about the matrix-exponential distributions used in the construction. In practice, we use a class of \emph{concentrated matrix-exponential distributions} found numerically in \citep{hht2020}, and for which there is relatively little known about their properties. This necessitates the generality of the convergence result. 
% \end{rem}

\section{Extension to arbitrary (but fixed) discretisation structures}
To conclude this chapter we include some remarks on how to extend the convergence results to arbitrary discretisation structures.

Throughout, we have assumed that all intervals are of width \(\Delta\), i.e.~\(|y_{\ell+1}-y_\ell|=\Delta\), and that on every interval the dynamics of the fluid queue are modelled based on the same matrix exponential representation \((\bs \alpha, \bs S, \bs s)\). These assumptions are, in fact, not necessary, but they do serve to simplify the presentation slightly. The convergence results can be extended to use different sequences of matrix exponential representations on each interval, provided that for each sequence of matrix exponential distributions, the variance tends to \(0\). Moreover, we can extend the results to intervals of arbitrary width, provided that the width of the intervals is not arbitrarily small. Here we describe how one would prove such results.

The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the Theorem depend only on the interval \(\calD_{\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval, and not on any other interval. Thus, Lemma~\ref{thm: a thm!} holds independently on each interval, as does Corollary~\ref{cor: aln222}.

Let the width of an interval \(\mathcal D_{\ell_0}\) be \(\Delta_{\ell_0}=y_{\ell_0+1}-y_{\ell_0}\) and suppose that sequence of matrix exponential random variables used to model the dynamics of the fluid queue on the interval \(\calD_{\ell_0}\) is \(Z_{\ell_0}^{(p)}\). Regarding the domination condition in Lemma~\ref{lem: another bound}, we can extend it the following version, 
\begin{lem}\label{lem: another bound sdfg}
	Assume \(\inf_{\ell_0}\Delta_{\ell_0}>0\) and \(\sup_{\ell_0}\var(Z_{\ell_0})<\infty\). Then, for all \(i\in\mathcal S_+\cup \calS_-\), \(\ell_0\in\mathcal K\setminus\{-1,K+1\},\) and \(n\geq 2\), 
	\begin{align}
		&\mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \Ydp(n-1) = (\ell_0,i), \tau_{n-1}^{(p)}\leq  E^\lambda ) \leq b_{\ell_0}^{(p)},\label{eqn: ell0 version}
	\end{align}
	where 
	\[b_{\ell_0}^{(p)} = 1-e^{-q(\Delta_{\ell_0}+\varepsilon_{\ell_0}^{(p)})}\left[1-e^{q\varepsilon_{\ell_0}^{(p)}-\lambda \Delta_{\ell_0}/|c_{min}|}\right] + \cfrac{\var(Z_{\ell_0}^{(p)})}{\left(\varepsilon^{(p)}\right)^2} + |r_{1,\ell_0}^{(p)}| \]
	and  
	\[|r_{1,\ell_0}^{(p)}|\leq 2G\cfrac{\var \left(Z_{\ell_0}^{(p)}\right)}{\left(\varepsilon_{\ell_0}^{(p)}\right)^2} + 2L\varepsilon_{\ell_0}^{(p)}.\]
	Hence, for all \(i\in\mathcal S_+\cup \calS_-\), and \(n\geq 2\), 
	\begin{align}
		&\mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \phi^{(p)}(\tau_{n-1}^{(p)})=i, \tau_{n-1}^{(p)}\leq  E^\lambda ) \leq b^{(p)}, \label{eqn: skjhg}
	\end{align}
	where 
	\[b^{(p)} = \max\left\{\sup_{\ell_0}b_{\ell_0}^{(p)}, \cfrac{q}{\lambda + q}\right\}.\]
\end{lem}
\begin{proof}
	For the proof of (\ref{eqn: ell0 version}) follow the same arguments as in the proof of Lemma~\ref{lem: another bound}. The bound in (\ref{eqn: skjhg}) follows by the assumptions in the statement of the result, since 
	\begin{align*}
		&\mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \phi^{(p)}(\tau_{n-1}^{(p)})=i, \tau_{n-1}^{(p)}\leq  E^\lambda ) 
		\\&\leq \sup_{\ell_0} \mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \Ydp(n-1) = (\ell_0, i), \tau_{n-1}^{(p)}\leq  E^\lambda ) 
		\\&\leq \max\left\{\sup_{\ell_0}b_{\ell_0}^{(p)}, \cfrac{q}{\lambda + q}\right\}.
	\end{align*}
\end{proof}
Given Lemma~\ref{lem: another bound sdfg}, then an equivalent of Lemma~\ref{lem: another bound 2} remains true, the proof of which follows verbatim except with the use of Lemma~\ref{lem: another bound} replaced by Lemma~\ref{lem: another bound sdfg}. Corollary~\ref{vcor: cdks d} remains true without modification. Lemma~\ref{lem: KajPOw} and Corollary~\ref{cor: lk} remain true without modification provided that \(\lim\limits_{p\to\infty}\var(Z_\ell^{(p)})\to 0\) for all \(\ell\).

\begin{rem}
	I suspect that the convergence results can also be extended to approximating so-called \emph{multi-layer fluid queues}, as described in \cite{bo2008}. 
\end{rem}