### 2022-02-10 appendix_math_background/appendix.tex

%!TEX root = ../thesis.tex
\chapter{More mathematical background}\label{app: math background}