### 2022-02-17 chapter6/chapter.tex

%!TEX root = ../thesis.tex
\chapter{Numerical investigations\label{sec: numerics}}
We have now established multiple schemes for approximating fluid queues which are suitable for approximating the performance measures of fluid-fluid queues derived in \cite{bo2014}. In this chapter we numerically investigate various aspects of the approximation schemes. Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin scheme (as described in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation scheme introduced by \cite{bo2013}. %Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume scheme with an upwind flux. 
Recall that, due to their stochastic interpretation, the uniformisation and QBD-RAP schemes are positivity preserving.

The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased. For simplicity, the number of basis functions is kept constant across all cells. We refer to the number of basis functions on a cell \(k\) as the \emph{dimension} and consider dimensions \(p=2r+1\) for \(r=0,1,2,3,...,10\). For discontinuous Galerkin scheme, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell. E.g.~if we use 3 basis functions in the discontinuous Galerkin scheme, we approximate the solution by a quadratic on each cell. For the QBD-RAP scheme, increasing the dimension of the scheme means increasing the order of the CME distribution used to construct the scheme. To make a comparable equivalent for the uniformisation scheme with a fixed cell size we divide each cell into smaller sub-cells over which we approximate the solution. I.e.~for a dimension \(p\) uniformisation scheme we divide each cell into \(p\) sub-cells. Equivalently, we may think of a dimension \(p\) uniformisation scheme as using \(p\) piecewise constant functions to approximate the solution on each cell. For all schemes (DG, QBD-RAP and uniformisation), if we construct an order \(p\) approximation, there are \(K\) cells, \(N\) phases and \(c_i\neq 0\) for all \(i\in\calS\), then the resulting approximation to the generator \(\mathbb B\) is a matrix of dimension \(pKN + N\). In this sense each approximation scheme leads to matrices of the same size (although not necessarily the same number of non-zero elements).

To keep the content of this chapter contained, we do not investigate all aspects of the schemes. For each numerical experiment we keep the cell size, \(\Delta\), fixed for the DG and QBD-RAP schemes. As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next \citep{nodalDGBook}. We investigate schemes with an upwind flux only. For schemes which require us to integrate over time we do not investigate the stability of the schemes with respect to the \(t\)-step-size, or the time-integration scheme itself (where required). Instead, we fix the time-integration step size for each numerical experiment at a suitably small value to obey a certain stability criterion (a CFL-like condition, \citep[Section~4.8]{nodalDGBook}). Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (\cite{sr2002}, \cite{nodalDGBook}~Section~5.7, see also Section~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.

For some numerical experiments we implement the discontinuous Galerkin scheme with and without a slope-limiter implemented (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}). The slope limiter avoids oscillations and negative solutions. However, we note that in the context of approximating the operator \(\mathbb\Psi\) for a fluid-fluid queue, it is not obvious how one might apply the concept of slope-limiting, other than to post-process the solution with a limiter or filter. Even when slope limiting is possible, the resultant approximation is at best linear around discontinuities. Moreover, there is a computational cost in applying a limiter, which could be significant, particularly if many initial conditions are to be considered. In contrast, for the uniformisation and QBD-RAP the approximate solution is guaranteed to produce positive probabilities without any post-processing or extra computation.

Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}). We investigate two schemes with slope limiting. In the first scheme, which we refer to the \emph{DG-lim} scheme, we use a DG scheme of dimension \(p\) and apply the Generalised MUSCL slope limiter as described in Section~\ref{sec: slope limiting}. In the second scheme, which we refer to as the \emph{DG-lin-lim} scheme, we follow the construction described in~\ref{sec: anoth pos pres}, which is effectively a DG scheme with cells of width \(2\Delta/(p+1)\), for \(p=1,3,5,...,21\), and a basis of dimension \(2\) (linear) on each cell, with an upwind flux, and we apply the Generalised MUSCL limiter. Thus, the DG-lin-lim scheme uses \(p+1\), \(p=1,3,5,...,21\), basis functions to represent the solution on each of the original cells \(\calD_k\), and we define the dimension of the scheme to be \(p+1\). We do not investigate filtering for the DG scheme.

The performance of the discontinuous Galerkin scheme has been well-studied in some contexts (\cite{c99}, \cite{nodalDGBook}~Section~5.5), and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions. Here, we mostly focus on investigating the numerical performance of the schemes on problems with non-smooth solutions. The purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP scheme. In the stochastic modelling community it is very common to have problems with discontinuities, such as a non-smooth initial conditions. As we will see (in Section~\ref{sec: transient approx}), even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities or non-smooth behaviour into the problem. A very specific set of conditions on the initial density and point masses must hold for the transient distribution to remain continuous as it evolves over time \citep{bo2014,bop2020}. Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues (see, for example, Section~\ref{sec: return approx}). Limiting distributions of fluid queues are smooth, however.

As we shall see, if the solution is smooth, then the DG scheme is highly-effective and exhibits rapid convergence to the solution as the number of polynomial basis functions is increased. However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin scheme, leading to nonsense solutions. In these cases, the QBD-RAP and DG-lin-lim approximations perform relatively well compared to the other positivity preserving schemes investigated.

Lastly, before we start the numerical investigation, note that all the following error plots are on a \(\log_{10}\)-\(\log_{10}\) scale. If the error function takes the form \(error = \beta_0^* p^{\beta_1}\), where \(p\) is the dimension of the scheme, then the \(\log_{10}\)-\(\log_{10}\) will show a linear trend; \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\), where \(\beta_0=\log_{10}(\beta_0^*)\). Where relevant, to estimate the asymptotic rate of convergence, \(\beta_1\), we estimate the of the line \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\) using ordinary least squares and the last eight data points on each plot. We plot the estimated lines of best fit and quote their slope to the right of the plots.

The structure of this chapter is as follows. In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different schemes and observe their performance. This allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem. In Section~\ref{sec: wave num} we investigate travelling wave problems with various initial conditions. For the travelling wave problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics. Next, we investigate a simple fluid queue with two phases. Section~\ref{sec:stat} investigates the ability of the schemes to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the model for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\). Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num} and approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\).

% THE STRUCTURE OF THE CHAPTER IS AS FOLLOWS.

% 1. FUNCTION APPROX./RECONSTRUCTION

% 2. 1d wave eqn

% 3. STATIONARY DISTRIBUTIONS
	% i. SIMPLER
	% ii. MORE COMPLEX

% 4. TRANSIENT DISTRIBUTIONS 
	% i. SIMPLER
	% ii. MORE COMPLEX

% 5. HITTING TIMES 
	% i. SIMPLER
	% ii. MORE COMPLEX

% 6. FLUID-FLUID QUEUES

% 7. CONCLUDING REMARKS

\section{Function approximation/reconstruction}\label{sec: recon num}
We start our investigation by looking at how well the schemes perform at approximating an initial condition. By approximating the initial condition only, we aim to instrument the performance of the approximation schemes without any dynamics.

For the discontinuous Galerkin method, we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}. For the purposes of approximation and reconstruction for the QBD-RAP scheme we must orientate the initial condition in a positive or negative direction; here, we suppose that the initial conditions belong to a positive phase. First we investigate three closing vectors we can use for the reconstruction in the QBD-RAP scheme. From the investigation, we decide to use the closing vector (\ref{eqn:5462}) throughout the rest of the chapter.

For this investigation we consider approximating initial conditions over a single cell of width \(\Delta = 1\). To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a trapezoidal rule with 10,001 function evaluations. Similarly, we use 10,001 function evaluations to approximate \(L^p\) errors, and also as a finite set of points over which to compute the KS statistics.

% First we briefly investigate the possible choices of closing operators as introduced in Section~\ref{sec: closing}.

\subsection{QBD-RAP closing operators}\label{sec: closing vecs numerics}
The three closing operators we investigate are the following. The unnormalised closing vector in Equation~(\ref{eqn: alal}). A normalised version of the closing operator in Equation~(\ref{eqn:546}), with the normalised version given by (\ref{eqn: nonlin}). We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector. Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme. The third closing vector we investigate is that in Equation~(\ref{eqn:5462}) and we refer to this as the \emph{normalised closing operator}.

\begin{example}\label{ex: another example}
	First consider approximating the initial condition with density function \(2\times 1(x<0.5)\). In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation with the three different closing vectors. Interestingly, in this case, the unnormalised closing operator outperforms the other two at low dimensions. In this case, it just so happens that the error due to truncation of the unnormalised reconstruction constructively interferes with other errors to cause the error to be lower at low dimensions for the unnormalised scheme.

	Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection. The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward). Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.

If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} (right), then we see that the unnormalised reconstruction performs the poorest. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests the naive normalised and normalised reconstruction are very similar. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: another example} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line). The error curves of the naive normalised (orange) and normalised closing vectors are almost coincident. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}%To estimate the asymptotic rate of convergence, we fit a linear model of \(\log_{10}(error) = \beta_0 + \beta_1\log_{10}(Dim)\) to the the last 6 data points on the curves above, which corresponds to a rate of convergence \(error \propto N^{-\beta_1}\). The estimate of \(\beta_1\) was approximately -1.0 in all cases.}
	\label{fig: fun 2 ks error qbdrap closing vecs}
\end{figure} 
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: al}
	Now consider approximating the initial density \(f(x)=1\). Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error and \(L^2\) error of the PDF we now see that the normalised reconstructions outperform the unnormalised reconstruction. This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: al} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line). The error curves of the error curves of the naive normalised (orange) and normalised closing vectors are almost coincident. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 4 ks error qbdrap closing vecs}
\end{figure}

Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for dimension 1, 3, 5 and 7, as well as the true PDF. Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots. This feature is due to a significant amount of mass being lost due to the truncation. In comparison, the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator. The `folding' in the normalised closing operator manifest itself as more mass at the left of these plots compared to the unnormalised reconstructions. % Recall that, as discussed in Section~\ref{}, for the reconstruction we must to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}. Here we suppose that the phase is positive, so the `folding' around \(\Delta\) in the normalised closing operators appears appears at the left of the plots. In general, reconstructions via the unnormalised closing operator typically underestimate the value of the function in this region.

Figure~\ref{fig: pdf reconstructed} also show that both closing operators fail to approximate the initial condition well at the right-hand side of the interval. Perhaps there is a different reconstruction method or another closing operator which could alleviate this issue. 
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
	\caption{Approximations to the PDF \(f(x)=1\) (orange dotted line) for Example~\ref{ex: al}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
	\label{fig: pdf reconstructed}
\end{figure} 
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: aasmm}
	Next we consider approximating the initial distribution with density \(-6x^2+6x\). Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable. However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the \(L^2\) error between the reconstructed PDF and the true PDF, then the unnormalised closing operator performs better than the two normalised ones for dimension 5 and above. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error for the PDFs (right) of Example~\ref{ex: aasmm}, for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line). The error curves for the naive normalised (orange) and normalised closing vectors are almost coincident. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 6 ks error qbdrap closing vecs}
\end{figure}

The fact that the unnormalised closing operator outperforms the two normalised ones can be explained by the `folding' of the normalised operators around \(\Delta\). In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\). In Figure~\ref{fig: pdf reconstructed quadratic} at the left of the plots, we observe that the normalised reconstructions over-estimate the density function, whereas the unnormalised reconstruction looks to be doing better. The `folding' in the normalised closing operator manifest itself as more mass at the left of these plots compared to the unnormalised reconstructions. 
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
	\caption{Approximations to the PDF \(f(x)=-6x^2+6x\) (orange dotted line) for Example~\ref{ex: aasmm}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).} 
	\label{fig: pdf reconstructed quadratic}
\end{figure} 
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: kjfk}
	Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\). This density function is at a maximum at the left of the region. Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case. Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that this is the case; the normalised reconstructions perform relatively well compared to the unnormalised reconstruction, as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF). If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: kjfk} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line). The error curves for the naive normalised (orange) and normalised closing vectors are almost coincident. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 7 ks error qbdrap closing vecs}
\end{figure}
\exampleFloatBarrier
\end{example}

In Table~\ref{table: error rates for QBD-RAP closing} we summarise the rates of convergence of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots). In general, the normalised closing operators converges at a rate approximately as fast, or faster, than the unnormalised reconstruction.
\begin{table}
	\centering
	\begin{tabular}{c|c|c|c}
			Example & PDF & \(\beta_1\), unnormalised & \(\beta_1\), normalised \\\hline
			\ref{ex: another example} & \(2\times 1(x<0.5)\) & -1.07 & -0.98 \\
			\ref{ex: al} & 1 & -0.99 & -0.91 \\
			\ref{ex: aasmm} & \(-6x^2+6x\) & -2.72 & -2.57 \\ 
			\ref{ex: kjfk} & \(3e^{-3x}/(1-e^-3)\) & -1.11 & -2.35 
	\end{tabular}
	\caption{\label{table: error rates for QBD-RAP closing} Estimated rates of convergence, \(\beta_1\), of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).}
\end{table}

In summary, the naive normalised and normalised closing operators perform almost identically. In some cases the unnormalised closing operator can outperform the normalised closing operators, typically when the mass of the function to be approximated at the left-hand edge of the cell is low. Further, the rates of convergence for all three schemes was similar, except for in Example~\ref{ex: kjfk} where we approximated the initial condition \(3e^{-3x}/(1-e^-3)\); in this case the rate of convergence of the normalised closing operators was approximately twice as large as the unnormalised version. Given the results above and its theoretical tractability, we will opt to use the normalised closing operator for the rest of the numerical experiments.

\FloatBarrier
\subsection{Comparison of schemes}\label{sec: comp}
Here we compare the ability of the QBD-RAP, uniformisation, and DG schemes to reconstruct initial conditions.

\begin{example}\label{ex: exa}
	First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5. This distribution does not have a PDF, so we compare the CDFs only. In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations. Observing the KS metric, it appears that none of the schemes converge and the KS error sits around 0.5 for all schemes. This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity and none of the schemes appear to converge at the discontinuity at \(x=0.5\). Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the schemes appear to show the convergent behaviour we expect. Here, the uniformisation scheme appears to perform the best, while the QBD-RAP scheme performs the worst. The rate of convergence of the QBD-RAP scheme appears to be similar to the rate of convergence of the DG scheme.

Perhaps it is no surprise that the uniformisation scheme performs best. In the uniformisation scheme as we increase the dimension we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution. As such, the uniformisation scheme can produce a piecewise continuous, linear approximation to the CDF. In contrast, both the DG and QBD-RAP schemes result in a smooth approximation to the CDF. Given the initial distribution is far from smooth (it's a point mass), then we might expect that the uniformisation scheme will perform relatively well. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^1\) error of the CDF (right) of Example~\ref{ex: exa} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 1 comp} 
\end{figure}

In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP schemes alongside the true CDF. The DG scheme displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1. On the other hand, although the QBD-RAP scheme converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1. The uniformisation scheme approximates the CDF exceptionally well.
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/comp/fun1/cdfs_formatted.pdf}
	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes for Example~\ref{ex: exa}. The true distribution function is \(1(x\geq 0.5)\) (grey dotted line).}
	\label{fig: pdf comp fun 1}
\end{figure} 
\exampleFloatBarrier
\end{example}

\begin{example}\label{eqn: example something}
	Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\). Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right). Figure~\ref{fig: fun 2 comp} suggests that all schemes converge at a similar rate for this problem. Here, the QBD-RAP scheme performs worst, the uniformisation scheme second, and the DG scheme the best. However, once again the DG scheme exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for Example~\ref{ex: exa}). 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{eqn: example something} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 2 comp} 
\end{figure}
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: alllll}
	So far we have considered two problems which exhibit discontinuities. At the other extreme we now consider an initial distribution with density \(-6x^2+6x\). In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs. Since the DG method projects the initial condition onto a polynomial basis, then, for a dimension 3 approximation and above, the DG scheme can approximate the initial condition exactly. This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from dimension 1 to 3, then plateaus. Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, the errors for the DG scheme are not 0. Regarding the other two schemes, they too appear to be convergent at approximately the same rate, with the uniformisation scheme performing better for the KS error, but very similarly in terms of the \(L^2\) error. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: alllll} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 6 comp} 
\end{figure}
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: aalllkkdd}Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\). Figure~\ref{fig: fun 9 comp} shows the KS and \(L^2\) errors for the PDF. Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG scheme approximates the initial condition very well. The uniformisation scheme performs second best, while the QBD-RAP scheme is slowest to converge. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.1cm 0.8cm 0.1cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.1cm 0.8cm 0.1cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: aalllkkdd} for the DG scheme (blue solid line), uniformisation scheme (orange dashed line) and QBD-RAP scheme (purple dashed line). The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 9 comp} 
\end{figure}
\exampleFloatBarrier
\end{example}

In conclusion, we have observed that the DG scheme performs well with respect to the error metrics considered, but can display oscillations and result in CDFs which are not non-decreasing. On the other hand, the uniformisation and QBD-RAP schemes result in non-decreasing approximations to the CDFs, but can converge slower with respect to the error metrics considered, with the QBD-RAP scheme often converging the slowest.
\FloatBarrier

\section{Travelling wave problems}\label{sec: wave num}
Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with various initial conditions. Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\). The PDE (if/when it exists) which describes this system is 
\[\cfrac{\partial}{\partial t} f(x,t) = -\cfrac{\partial}{\partial x} f(x,t),\]
where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
\[f(x,t) = f(x-t,0)\]
so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\). This example is convenient as it has a known solution and no stochastic dynamics, hence we can instrument the ability of the schemes to approximate the flow of mass, without any stochastic dynamics.

We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space and discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric. Furthermore, we use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG scheme we also implement the Generalised MUSCL slope limiter in the form of the DG-lim and DG-lin-lim schemes (see also Section~\ref{sec: slope limiting}).

To investigate the performance of the schemes without the need to reconstruct the function within each cell we use the cell-wise error metric obtained by computing 
\begin{align}
	&\sum_{\ell=1}^{10} \left|\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,\ell,1)\right|\nonumber 
	\\&+\left|\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,11,1)\right| \label{eqn: cell errors}
\end{align}
where 
\begin{align}
	\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs boundary}
\end{align}
is the mass at the boundary, \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,11,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).

\begin{example}\label{ex: wave 4}
First consider the initial condition with PDF \(1(x<1)\). The level of the fluid queue is uniformly distributed over the first cell. For the DG-based schemes (including the uniformisation scheme) the initial condition can be represented exactly whereas, for the QBD-RAP scheme, it cannot. Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation schemes. Moreover, at time \(t=4\), the solution is \(f(x,4)=1(x\in[4,5))\). The projections related to the DG-based schemes can represent this solution exactly too, hence we might expect these schemes to work well here.

We plot the cell-wise error metric in Figure~\ref{fig: fun 4 wave cp} and observe that the DG scheme appears to converge fastest, followed by the QBD-RAP and DG-lin-lim schemes which perform comparably, then the uniformisation scheme. The worst performer is the DG-lim scheme which does not appear to converge. The DG scheme with a slope limiter has detected oscillations in the approximate solution and reduced the scheme to linear where oscillations are detected, thereby decreasing the accuracy of the scheme to linear where oscillations are present. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/L1_cell_probs.pdf}
	\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave in Example~\ref{ex: wave 4}. Plotted are the cell-wise errors for the DG (blue solid line), DG-lim scheme (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim scheme (gold dashed line) schemes, versus the dimension of the approximation. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: fun 4 wave cp} 
\end{figure}

Figure~\ref{fig: fun 4 wave} plots the KS error between the approximated and true CDFs and \(L^2\) error between the approximated and true PDFs. Comparing Figure~\ref{fig: fun 4 wave cp} with Figure~\ref{fig: fun 4 wave}, each scheme seems to show a similar rate of decay for these two error metrics as they did for the cell-wise errors.

Continuing with Figure~\ref{fig: fun 4 wave}. The DG-lim scheme does not appear to converge with these error metrics. Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to be converging with the QBD-RAP and  DG-lin-lim schemes converging similarly and faster than the uniformisation scheme. With these error metrics, the DG scheme is converging fastest. However, if we observe the approximations resulting from the DG scheme (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscillations and negative values.

Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP schemes for various dimension schemes. In the first row we observe that the DG scheme (sans limiter) produces an approximation to the PDF with negative values when the dimension is greater than 1. In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter, the DG approximation does not change significantly after dimension 3. This is due the fact that the DG scheme is at best linear in the presence of oscillations. In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation scheme, in the fourth row is the solution approximated using the QBD-RAP scheme, and in the last row is the approximation from the DG-lin-lim scheme.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 4} using the approximations from the DG (blue solid line), DG-lim scheme (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: fun 4 wave} 
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row) and DG-lin-lim schemes, for dimension 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}. The true density function is \(1(4\leq x<5)\).}
	\label{fig: pdf wave fun 4}
\end{figure}

This is a particularly interesting example. It shows that, for the DG scheme, even though there is no discretisation error for the initial condition, we use a strong stability preserving time-integration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}. 
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: wave 2}
Another interesting example occurs with the initial condition with CDF \(1(x\geq 0.5)\), i.e.~a point mass at 0.5. The exact solution at time \(t=4\) is therefore a point mass at \(4.5\). No PDF exists for the true distribution, so instead we compare the CDFs. Also recall that, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative due to the lack of point-wise convergence at the discontinuity, so for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs) instead, and also with the cell-wise error from (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}).

Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs. The \(L^1\) metric tells a similar story to the previous analysis: the DG-lim scheme does not converge as dimension increases, the other three positivity preserving schemes (the uniformisation, DG-lin-lim and QBD-RAP schemes) appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the uniformisation scheme. The DG scheme performs the best. However, if we plot the approximations to the CDFs from the DG scheme (not shown) we would once again see an oscillating, non-monotonic function.

Another interesting observation is to compare the performance of the uniformisation and QBD-RAP schemes with respect to the \(L^1\) metric on the CDFs in Figure~\ref{fig: fun 2 wave} (left) to that in Figure~\ref{fig: fun 1 comp} (right). Recall, in Figure~\ref{fig: fun 1 comp} we investigated the ability of the approximation schemes to represent the initial distribution with CDF \(1(x\geq 0.5)\), which is the initial condition of the problem we are considering here. In Figure~\ref{fig: fun 1 comp} (right) the uniformisation scheme out-performed the QBD-RAP scheme at reconstructing the initial condition. However, in Figure~\ref{fig: fun 2 wave} (left) we see that the QBD-RAP scheme out-performs the uniformisation scheme. This suggests that the QBD-RAP scheme is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation scheme.

Figure~\ref{fig: fun 2 wave} (right) plots the cell-based error metric in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}). Once again, the DG-lim scheme does not appear to converge. Interestingly, the error curve for the DG scheme is not monotonic. Furthermore, the DG scheme can result in negative estimates of the cell probabilities in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (plots not shown). Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme all appear to converge, with the QBD-RAP and DG-lin-lim schemes perform similarly and better than the DG scheme when the dimension is sufficiently large. The estimated rate of convergence of the QBD-RAP and DG-lin-lim schemes is approximately -4 for this example, which is significantly faster than the rate of convergence of these schemes in all other examples in this section (typically around -2 to -1). 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/meshs_l1_cdf_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/L1_cell_probs.pdf} 
	\caption{\(L^1\) error of CDFs (left) and the cell-wise error metric in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (right) for the travelling wave problem in Example~\ref{ex: wave 2} where approximation were constructed via the DG (blue solid line), DG-lin-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: fun 2 wave} 
\end{figure}
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: wave 6}
	Now consider an initial distribution which is truncated Gaussian with mean parameter 2.5 and standard deviation parameter 0.5, and truncated at the boundaries, 0 and 10;
	\begin{align}
		\mu([0,x)) := \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)1(0 \leq x < 10.0)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10\leq x), \label{eqn: kdjfksdf}
	\end{align}
	where \(\Phi(x)\) is the CDF of the standard normal distribution.

	At time \(t=4\) the distribution is truncated Gaussian with mean parameter 6.5, standard deviation parameter 0.5, and is truncated below at 4.0 and above at 10, and there is also a small mass at the upper boundary; 
	\begin{align}
		&\mathbb P(X(4)\leq x, \varphi(4)=1\mid X(0)\sim\mu, \varphi(0)=1)  
		=\cfrac{\Phi((x-6.5)/0.5)1(4.0\leq x)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10 \leq x).\label{eqn: bbbbbaabbaa}%\cfrac{\Phi(7.5/0.5)-\Phi(3.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}.
	\end{align}
	At time \(t=4\) the mass at the boundary is approximately \(1.28\times 10^{-12}\) and there is a small discontinuity at \(x=4\) where the CDF jumps from \(0\) to \(\sim 2.87\times 10^{-7}\), due to the truncation of the initial distribution at \(0\).

	Since the discontinuity at \(t=4,\,x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG scheme will perform well for this example. Figures~\ref{fig: fun 6 wave cp} and~\ref{fig: fun 6 wave} confirms that this is indeed the case. For all three error metrics the error obtained by the DG scheme rapidly decreases to a point where it is swamped by other numerical errors. This is characteristic of the DG scheme for the smooth problems we investigate throughout this chapter. For low order DG schemes there are regions where the approximated PDF is negative, however, as the order of the DG scheme increases, these regions disappear.

	Interestingly, even though this is a relatively smooth problem, the DG-lim scheme does not perform well. It must be that the initial and transient distributions are `sufficiently pointy' that oscillations in the numerical solutions occur and the limiter reduces the order of the scheme to linear.

	The other three positivity preserving schemes appear to converge, with the uniformisation scheme performing the worst and the DG-lin-lim scheme performing the best of the three for all error metrics. For the KS and \(L^2\) error on the PDFs in Figure~\ref{fig: fun 6 wave} the DG-lin-lim and QBD-RAP schemes perform similarly, however, for the cell-wise error in Figures~\ref{fig: fun 6 wave cp} the DG-lin-lim scheme performs significantly better than the QBD-RAP scheme. Given that the DG-lin-lim scheme performed similarly to the QBD-RAP scheme for the previous examples, and also for the other error metrics for this example, this is unexpected.

	To investigate the performance of the QBD-RAP and DG-lin-lim schemes, in Figure~\ref{fig: fun 6 wave pdfs} we plot the dimensions 21 QBD-RAP and dimensions 22 DG-lin-lim approximations to the PDF. Figure~\ref{fig: fun 6 wave pdfs} shows that the DG-lin-lim scheme approximates the PDF well, except for at the peak located at \(x=6.5\) (I suspect the limiter has affected the solution in this region). Specifically, the DG-lin-lim scheme underestimates the density in the peak, but overestimates the density either side of the peak. As a result, when we integrate over \(x\in[6,7]\) to compute the cell-wise probabilities, errors cancel over this region and the DG-lin-lim scheme produces an accurate estimate of the probability in this cell. In comparison, the QBD-RAP scheme also underestimates the density at the peak, but also within the entire interval \([6,7]\), albeit less so at the edges than the middle. The QBD-RAP scheme also overestimates the density in the adjacent cells, \([5,7]\) and \([7,8]\). Hence, the overall cell-wise error is large. 
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/L1_cell_probs.pdf}
		\caption{Cell-wise error defined in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) for the travelling wave problem in Example~\ref{ex: wave 6}. Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}  
		\label{fig: fun 6 wave cp} 
	\end{figure}
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_ks_error_formatted.pdf}%
		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_l2_pdf_error_formatted.pdf}
		\caption{KS error (left) and \(L^2\) error for the PDFs (right) for Example~\ref{ex: wave 6} where approximations were constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
		\label{fig: fun 6 wave} 
	\end{figure}
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun6/order_21_PDFs_formatted.pdf}
		\caption{Dimension 21 QBD-RAP (purple) and dimension 22 DG-lin-lim (gold) approximations to the transient PDF at time \(t=4\) for Example~\ref{ex: wave 6}. The true PDF is plotted in grey.}  
		\label{fig: fun 6 wave pdfs} 
	\end{figure}
\exampleFloatBarrier
\end{example}

\begin{example}\label{ex: wave 1}
	We now want to look at how the schemes might handle a mass at the boundary. We introduce an ephemeral second phase into the model with phase transition rate \(T_{22}=-1\), and fluid rate \(c_2=0\). The generator is therefore 
	\[T=\left[\begin{array}{cc} 0 & 0 \\ 1 & -1 \end{array}\right].\]

	We suppose that the initial condition is a point mass at the boundary in phase \(2\), i.e. \[\mathbb P(X(0)\leq x, \varphi(0)=2)=1(0\leq x).\] 
	With this initial condition, the transient distribution at time \(t=4\) is 
	\begin{align}
		&\mathbb P(X(4)\leq x,\varphi(4)=1 \mid X(0)=0,\varphi(0)=2) \nonumber 
		\\&= e^{T_{22}4}\left(e^{-T_{22}x}-1\right)1(4< x) + (1-e^{T_{22}4})1(4\geq x) \nonumber
		\\&= e^{-4}\left(e^{4x}-1\right)1(4< x) + (1-e^{-4})1(4\geq x) \label{eqn: asjda}
	\end{align}
	and 
	\begin{align}
		\mathbb P(X(4)\leq x,\varphi(4)=2 \mid X(0)=0,\varphi(0)=2) = e^{-4}1(0\leq x).
	\end{align}
	The PDF at \(t=4\) is discontinuous at \(x=4\), which is at the edge of a cell. For this problem, all schemes can represent the initial condition exactly.

	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (left) and the \(L^2\) error between the PDFs (right). Observing Figures~\ref{fig: fun 1 wave cp} and~\ref{fig: fun 1 wave}, the DG-lim scheme does not perform well, which might be expected due to the discontinuity in the transient distribution at \(x=t<10\) in Phase~1. The uniformisation, DG-lin-lim and QBD-RAP schemes appear to converge, with the QBD-RAP scheme converging fastest, but performing similarly to the DG-lin-lim scheme, and the uniformisation scheme performs the poorest of the three. The DG scheme performs the best, however, produces approximations which are negative and oscillatory as we might expect given the discontinuity. A selection of approximations to the transient PDF are shown in Figure~\ref{fig: pdf wave fun 1}.
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/L1_cell_probs.pdf} 
		\caption{Cell-wise error for the travelling wave problem in Example~\ref{ex: wave 1}. Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}  
		\label{fig: fun 1 wave cp}  
	\end{figure}
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_ks_error_formatted.pdf}%
		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
		\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 1} where approximations are constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
		\label{fig: fun 1 wave} 
	\end{figure}
	\begin{figure}[h]
		\centering
		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun1/pdf_formatted.pdf}
		\caption{Approximate transient PDFs at time \(t=4\) for Example~\ref{ex: wave 1} using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row), and DG-lin-lim schemes of dimensions 1, 3, 5, and 7 (columns). The true density function is \(e^{-4}e^{x}1(x<4)\).} 
		\label{fig: pdf wave fun 1}
	\end{figure} 
	\exampleFloatBarrier
\end{example}

In summary, we applied the DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes to travelling wave problems with various initial conditions. All the problems induced oscillations in the DG scheme at some point and as a result the DG-lim scheme reduced the order of the approximation to linear. In some cases, the oscillations in the DG scheme appear to be transient and were not material in the final approximation. In other cases, the oscillations were present in the final DG approximation causing the approximation to take unphysical values. The DG-lin-lim, uniformisation and QBD-RAP schemes are three positivity preserving alternatives. In general, the uniformisation scheme performs worse than the DG-lin-lim and QBD-RAP scheme, and appears to converge slowly to the true solution. The DG-lin-lim and QBD-RAP schemes perform similarly with the DG-lin-lim scheme performing slightly better for the smoother problems and the QBD-RAP scheme performing slightly better for the problems with more significant discontinuities and point masses.

\FloatBarrier
\section{Stationary distributions} \label{sec:stat}
We briefly turn our attention to stationary distributions of fluid queues. Since the stationary distribution is smooth we expect that the DG scheme will work well here. The stationary distribution is convenient for numerical experiments as we can evaluate it analytically \citep{s2017}, and it does not require us to approximate initial conditions or integrate over time. Hence, analysing the performance of the approximation schemes with the stationary distribution allows us to instrument the ability of the schemes to capture the stochastic dynamics, without numerical integration in time.

Here we analyse a simple model which is based on Example~2 in \cite{bean2009}, except here we add a lower boundary to the model (no lower boundary is specified in Example~2 of \cite{bean2009} as it is inconsequential to their analysis).

% \begin{model}\label{model: simple}
% 	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2,3,4\}\), generator 
% 	\[\bs T = \left[\begin{array}{cccc}
% 		-1.1 & 1.1 & 0 & 0 \\
% 		1 & -1 & 0 & 0 \\ 
% 		0.01 & 0 & -0.01 & 0 \\
% 		0 & 0.01 & 0 & -0.01 
% 	\end{array}\right],\]
% 	and there are associated rates \(c_1=1, c_2 = -1, c_3=0, c_4=0\) and boundaries at \(x=0\) and \(x=10\). We specify two types of behaviour at the boundary.
% 	\begin{enumerate}[a)]
% 		\item (Absorbing model) Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
% 		\[\vligne{p_{21} & p_{22} & p_{23} & p_{24}} = \vligne{0 & 0 & 1 & 0}.\]
% 		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
% 		\[\vligne{p_{11} & p_{12} & p_{13} & p_{14}} = \vligne{0 & 0 & 0 & 1}.\]
% 		\item (Reflecting model) Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
% 		\[\vligne{p_{21} & p_{22} & p_{23} & p_{24}} = \vligne{1 & 0 & 0 & 0}.\]
% 		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
% 		\[\vligne{p_{11} & p_{12} & p_{13} & p_{14}} = \vligne{0 & 1 & 0 & 0}.\]
% 	\end{enumerate}
% \end{model}
% Phases~3 and 4 are included only to capture behaviour at the boundary.

% For Model~\ref{model: simple}~a), upon hitting the boundary, the process stays at the boundary for an exponentially distributed amount of time with mean 100. For Model~\ref{model: simple}~b), upon hitting the boundary, the process is immediately reflected. In Model~\ref{model: simple}~b) phases 3 and 4 are ephemeral and are included only for notational convenience.

\begin{model}\label{model: simple}
	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2,3,4\}\), generator 
	\[\bs T = \left[\begin{array}{cccc}
		-1.1 & 1.1 \\
		1 & -1 \\ 
	\end{array}\right],\]
	and there are associated rates \(c_1=1, c_2 = -1,\) and boundaries at \(x=0\) and \(x=10\). We specify two types of behaviour at the boundary.

	Upon hitting the lower boundary, the process transitions from Phase \(2\) to Phase \(j\) with probability \(p_{2j}\) where 
		\[\vligne{p_{21} & p_{22} } = \vligne{1 & 0}.\]
		Upon hitting the upper boundary, the process transitions from Phase \(1\) to Phase \(j\) with probability \(p_{1j}\) where 
		\[\vligne{p_{11} & p_{12} } = \vligne{0 & 1}.\]
	Thus, upon hitting the boundary, the fluid queue immediately transitions to the other phase and is reflected.
\end{model}

The model was discretised using the DG, uniformisation and QBD-RAP schemes using ten cells of width \(1\). We compute the coefficients for the stationary distribution in the following way. Suppose that a given discretisation scheme results in an approximation to the generator of the fluid queue as a matrix \(\bs B\). Then the stationary coefficients are found by solving 
\begin{align}
	\bs b \bs B = 0,\\
	\mbox{such that } \bs b\bs 1=1,
\end{align}
for the coefficients \(\bs b\).

% \paragraph{Model~\ref{model: simple}~a) Absorbing model}
% Figure~\ref{fig: absorbing stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations. Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors. The QBD-RAP scheme and uniformisation scheme both appear to be converging, with the errors for the QBD-RAP scheme decreasing faster.
% \begin{figure}[h]
% 	\centering
% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/ks_error_formatted.pdf}%
% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/l2_pdf_error_formatted.pdf}
% 	\caption{KS error between the true stationary CDF  and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~a) using the DG scheme (blue solid line), uniformisation scheme (orange dashed line) and QBD-RAP scheme (green dash-dotted line).} 
% 	\label{fig: absorbing stationary} 
% \end{figure}

% \paragraph{Model~\ref{model: simple}.}
Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations and the \(L^2\) error between the true stationary PDF and the approximations. Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors. The QBD-RAP and uniformisation schemes both appear to be converging, with the errors for the QBD-RAP scheme decreasing faster.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/l2_pdf_error_formatted.pdf}
	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Model~\ref{model: simple} using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dotted line) schemes. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: reflecting stationary} 
\end{figure}

\FloatBarrier
\section{Transient distributions}\label{sec: transient approx}
Once again we consider Model~\ref{model: simple} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells with width \(\Delta=1\)). Two initial conditions are considered, a point mass at 0 in Phase~1, and the initial distribution with PDF 
\begin{align}
	\cfrac{1}{2}e^{-x}/(1-e^{-10})\label{eqn: exp init cond}
\end{align}
in phases 1 and 2, with no mass at the boundaries. We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG without a limiter and also the DG-lim and DG-lin-lim schemes which utilise a Generalised MUSCL slope limiter.

To obtain a \emph{ground truth} 5,000,000 realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the masses within each cell were computed from the simulations. We do not attempt to numerically approximate the true PDF via simulation. We then compute the KS and \(L^1\) error metrics between the approximated and simulated CDFs, as well as the cell-wise error metric, given as follows. 
\begin{align}
	&\sum_{j\in\{1,2\}}\sum_{\ell=1}^{10} \left|\mathbb P_{sim}(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0)=0.5,\varphi(0)=1)-p(2,\ell,j)\right|\nonumber 
	\\&+\left|\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))-p(2,11,4)\right|\nonumber 
	\\&+\left|\mathbb P_{sim}X(2)\in\{0\}, \varphi(2)=3\mid X(0),\varphi(0))-p(2,0,3)\right| \label{eqn: cell errors 2}
\end{align}
where \(p(2,\ell,j)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0),\varphi(0))\), \(p(2,11,4)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))\), and \(p(2,0,3)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=3\mid X(0),\varphi(0))\), and \(\mathbb P_{sim}\) denotes an empirical probability evaluated using the simulations of the fluid queue.

To account for possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples. For the bootstrap we sample 5,000,000 realisations of the fluid queue with replacement from the original 5,000,000 samples, then compute error metrics with the resampled data. We resample 1,000 times, resulting in 1,000 estimates of the errors. Via the bootstrap, we report the \(5\)th and \(95\)th percentile of the sampling distribution of the errors.

To evaluate error metrics, we use a grid of 10,001 evenly spaced points for each phase.

To approximate the point mass initial condition we compute the initial coefficients for each scheme exactly. For the exponential initial condition in (\ref{eqn: exp init cond}) we compute the initial coefficients via Gauss-Lobatto quadrature for the DG scheme, by using the mid-point rule for the uniformisation scheme, and by using a trapezoidal rule with 2,001 points on each cell for the QBD-RAP scheme.

\paragraph{Model~\ref{model: simple} with exponential initial condition}
Figure~\ref{fig: reflecting transient exp} shows the KS and \(L^1\) metric on the CDF  for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes). For both error metrics the DG scheme converges rapidly until computational errors become significant. The uniformisation, DG-lin-lim and QBD-RAP schemes converge at a slower rate than the DG scheme, with the QBD-RAP and DG-lin-lim schemes converging faster than the uniformisation scheme. For the KS metric, the DG-lin-lim scheme outperforms the QBD-RAP scheme at all dimensions, while for the \(L^1\) metric, the two schemes perform similarly.

The DG-lim does not appear to be converging, which suggests there is at least one iteration during the numerical integration over time at which the numerical solution displays oscillations. However, when the DG approximations for the distribution at time \(t=2\) are plotted, they do not appear to display oscillations (not shown). This suggests that the oscillations which might occur with the DG scheme are transient, and have dissipated by time \(t=2\). The oscillations which the limiter detects in the DG scheme are likely from the reflecting boundary at \(x=0\). In the early stages of the evolution of the model the majority of the density in Phase~2 near zero will move to the left at rate 1 and hit the boundary at \(x=0\). Upon hitting the boundary, the density is reflected into Phase~1. Meanwhile, the initial density in Phase~1 moves to the right at rate 1 and the density reflected at the boundary from Phase~2 fills in to the region between \(x=0\) and \(x=t\) in Phase~1. Intuitively, in the early evolution of the model, there will be a sharp peak in the transient density at \(x=t\). I suspect that the true transient density is continuous, but not differentiable, at the point \(x=t\) for \(t<10\).  %
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: reflecting transient exp} 
\end{figure}

Figure~\ref{fig: reflecting transient exp cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition. Since the cell-wise errors do not require us to reconstruct the value of the solution within each cell, then this metric allows us to observe the error characteristics of the schemes without reconstruction. Figure~\ref{fig: reflecting transient exp cp} shows similar convergence characteristics to the KS error metric in Figure~\ref{fig: reflecting transient exp} (left). 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/L1_cell_probs_error_formatted.pdf}
	\caption{Cell-wise error metric from (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: reflecting transient exp cp} 
\end{figure}
\exampleFloatBarrier
\paragraph{Model~\ref{model: simple} with a point-mass initial condition}
Figure~\ref{fig: reflecting transient pm} shows the KS and \(L^1\) metric on the CDF  for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes). Comparing the error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition all schemes perform worse for the point mass initial condition. Regarding comparative convergence of error for this problem, for both error metrics the DG scheme converges fastest, followed by the QBD-RAP and DG-lin-lim schemes which converge comparatively, then the uniformisation scheme, while the DG-lim does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity. %
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}
	\caption{KS (left) and \(L^1\) (right) errors between the CDF at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: reflecting transient pm} 
\end{figure}

Figure~\ref{fig: reflecting transient pm cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the point mass initial condition. Figure~\ref{fig: reflecting transient pm cp} suggests that approximating the cell-wise error seems to be a difficult problem. This is likely caused by the discontinuity at \(x=2\) in Phase \(1\), which lies exactly on a cell boundary. To investigate how the position of the discontinuity might affect the error metrics we evolved the model to time \(t=2.1\) and computed the KS error, \(L^1\) error between the CDFs and the cell-wise error. Once again, we use simulation and bootstrapping to approximate the true distribution. The plots (not shown) of KS error and \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) are relatively similar and not noteworthy. The plot of the cell-wise error metric at time \(t=2.1\) in Figure~\ref{fig: reflecting transient pm t2.1} is somewhat more interesting. Firstly, Figure~\ref{fig: reflecting transient pm t2.1} shows that all schemes improved from time \(t=2\) to time \(t=2.1\) with respect to this metric. Figure~\ref{fig: reflecting transient pm t2.1} also shows that the cell-wise error metric is somewhat volatile for the DG scheme; I suspect that this is due to the oscillations in the DG approximation. In contrast, in Figure~\ref{fig: reflecting transient pm t2.1} the uniformisation and QBD-RAP schemes have monotonically decreasing error curves. Also interesting is the DG-lin-lim scheme, which does not improve past dimension 7. I suspect that the discontinuity at \(x=t\) is not easily resolved by the scheme given the presence of the limiter. Figure~\ref{fig: reflecting transient pm t2.1 cdfs} plots the CDFs in Phase~1 around the discontinuity in the transient distribution at time \(t=2.1\) for the dimension 21 QBD-RAP and dimension 22 DG-lin-lim schemes. Neither scheme does a particularly good job of capturing the discontinuity, but the QBD-RAP scheme appears to perform slightly better. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/L1_cell_probs_error_formatted.pdf}
	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2\) with the point mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.} 
	\label{fig: reflecting transient pm cp} 
\end{figure}%
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1/L1_cell_probs_error_formatted.pdf}
	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2.1\) with the point mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: reflecting transient pm t2.1} 
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1/order_21_CDFs_formatted.pdf}
	\caption{Dimension 21 QBD-RAP and dimension 22 DG-lin-lim approximations to the CDF in Phase~1 for Model~\ref{model: simple} with a point-mass initial condition. A CDF estimated via simulation is plotted in grey.}  
	\label{fig: reflecting transient pm t2.1 cdfs} 
\end{figure}

In summary, for the exponential initial condition the DG scheme performed well, producing the lowest errors. This was followed by the DG-lin-lim scheme, then the QBD-RAP scheme, then the uniformisation scheme, while the DG-lim scheme did not perform well at all. The fact that the DG-lim scheme did not perform well suggests that there must have been oscillations in the DG solution at some point, although these oscillations are not obvious in the DG approximation at time \(t=2\). In general the methods all performed better for the exponential initial condition than for the point mass initial condition. For the point mass initial condition the DG scheme produced oscillatory solutions. Of the positivity preserving methods, the DG-lim scheme performed poorly and did not appear to be converging, the uniformisation scheme appears to be converging but at a slower rate than the DG-lin-lim and QBD-RAP schemes. Of the two most performant positivity preserving schemes, the QBD-RAP scheme performed slightly better for the point mass initial condition, while the DG-lin-lim scheme performed slightly better for the exponential initial condition.

\FloatBarrier
\section{Hitting times}\label{sec: return approx}
Once again we consider Model~\ref{model: simple}. Let \(\zeta_{X}(\{0,1\}) = \{\inf t>0 \mid X(t)=0, \mbox{ or }X(t)=1\},\) be the first hitting time of \(\{X(t)\}\) on the set \(\{0,1\}\). The distribution of the hitting time in phase \(i\in\{1,2\}\) is 
\begin{equation}\label{eqn: hit cdf}\mathbb P(\zeta_{X}(\{0,1\}) < t, \varphi(t)=i\mid \bs X(0)\sim \mu),\end{equation}
for some initial distribution \(\mu\). We look at two initial conditions; an exponential with equal mass in each phase, 
\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = \cfrac{1}{2}\cfrac{e^{-x}}{(1-e^{-1}},\,i\in\{1,2\},\]
and a point mass at \(X(0)=0\) in phase \(\varphi(0)=1\).

We use the DG, uniformisation and QBD-RAP schemes to discretise the fluid queue and partition \([0,1]\) into three intervals of width \(1/3\). To capture the mass which has hit the set \(\{0,1\}\), we suppose that when the process hits \(\{0\}\) or \(\{1\}\) it is absorbed forever and remains in the phase in which the process first hit \(\{0\}\) or \(\{1\}\).

We integrate the schemes until time \(t=10\) using the SSPRK4 method with \(t\)-step size 0.005/3.\footnote{Since we use a smaller cell-width in this example than in previous examples we need to reduce the \(t\)-step size accordingly to ensure that numerical integration is stable for schemes up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We use the DG scheme with and without a slope limiter during the time-integration and implement both the DG-lim and DG-lin-lim schemes. At each time-step of the numerical integration, we record the amount of mass at the absorbing boundaries in each phase, which gives us an approximation of the cumulative distribution function of the hitting time in each phase up to time \(t=10\).

As a \emph{ground truth} we simulated 5,000,000 realisations and recorded the hitting time on the set \(\{0,1\}\) and the phase at the time of hitting. We then compute the empirical CDF of the hitting probabilities 
\[\mathbb P(\zeta_{X}(\{0,1\}) < t, \varphi(t)=i\mid \bs X(0)\sim \mu),\] 
for \(t=0.005/3\times k\), \(k=0,...,6000\).

To account for Monte-Carlo errors we took 1,000 bootstrap resamples of the original 5,000,000 samples and computed the empirical CDF of the hitting probabilities for each bootstrap sample. For each bootstrap sample, we resampled 5,000,000 points with replacement from the original 5,000,000 realisations, then computed the error metrics between the empirical and approximated CDFs and, ultimately, estimated the 5th and 95th percentile of the distribution of the errors.

\paragraph{Exponential initial condition}
Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the five numerical approximation schemes applied to the hitting time problem with the exponential initial condition. For both error metrics the DG scheme performs the best. For the \(L^1\) error metric the DG scheme converges up to dimension 3, after which there is little improvement in the error; I believe that this is due to other numerical errors dominating. The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation. Intuitively, with this initial condition we suspect that the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x,t<1\). We also expect there is a discontinuity in the hitting time PDF at time \(t=1\). Regarding the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and converging at a faster rate than the uniformisation scheme.

In Figure~\ref{fig: hitting time oscillation exp} we plot the PDFs of the hitting times \(\zeta_X(\{0,1\})\), in each phase, for the dimension 21 DG, dimension 21, QBD-RAP and dimension 22 DG-lin-lim schemes. Figure~\ref{fig: hitting time oscillation exp} demonstrates oscillations in the DG approximation, though the PDF is not negative in this case. All schemes seem to capture the discontinuity at \(t=1\) relatively well. There is an interesting artefact in the QBD-RAP approximation around \(t=0\) where the scheme has generated an oscillation. I suspect that this artefact should disappear as the dimension of the scheme is increased, unlike the DG scheme where oscillations can persist when the dimension is increased. Let's investigate this artefact further.

Intuitively, most sample paths which exit near \(t=0\) start near the boundary then see a change of phase shortly after \(t=0\) and remain in that phase until hitting \(\{0,1\}\). Consider such a sample path which starts at \(x_0=0\) in Phase~1, then exits at time \(v\) with a single change of phase at time \(v-u>0\). The QBD-RAP approximation to this sample paths has density\footnote{One advantage of the QBD-RAP and uniformisation approaches is that, from their stochastic interpretation, we can use sample-path arguments to analyse the scheme.} 
\begin{align}
	\bs \alpha e^{(\bs S - \bs I)(v-u)} \bs D e^{(\bs S - 1.1\bs I)u}\bs s = \bs \alpha e^{\bs S(v-u)} \bs D e^{\bs Su}\bs s e^{-v-0.1u},\label{eqn :lf}
\end{align}
and since the sample path will hit \(\{0\}\) at \(v=2u\), intuitively, we would expect (\ref{eqn :lf}) to approximate a point mass at \(v=2u\). To see that this is the case, recall that we can approximate a point mass at \((v-u)\) by 
\begin{equation}\bs k(\Delta - (v-u))\bs e^{\bs Sx}\bs s,\label{eqn:aafba}\end{equation}
where \(\bs k(z)=\cfrac{\bs \alpha e^{\bs Sz}}{\bs \alpha e^{\bs Sz\bs e}}\),
and recall that the matrix \(\bs D\) is an approximation; 
\begin{align}
	\bs \alpha e^{\bs S(v-u)}\bs D &= \mathbb E\left[\bs k(Z-(v-u))1(Z>v-u)\right]\nonumber 
	% \\&\approx \bs \alpha e^{\bs S(v-u)}\bs e \bs k(\Delta-(v-u))
	\\&\approx\mathbb P(Z>v-u) \bs k(\Delta-(v-u)), \label{eqn: aegpb}
\end{align} where \(Z\sim ME(\bs \alpha,\bs S)\)
and \(Z\) is concentrated around \(\Delta\).

Thus, (\ref{eqn :lf}) is approximately  
\[e^{-v-0.1u} \mathbb P(Z>v-u) \bs k(\Delta-(v-u))\bs e^{\bs Su}\bs s,\]
which approximates a point mass at \(u=v-u\), or \(2u=v\). Retracing our arguments, sources of error in this approximation come from the approximation in (\ref{eqn: aegpb}) and from the approximation of a point mass by (\ref{eqn:aafba}), which we commented on in Section~\ref{sec: closing vecs numerics} where we conducted numerical experiments about the unnormalised closing operator given by \(\bs v(x) = e^{\bs S x}\bs s\). Indeed, we saw that the unnormalised closing operator performed poorly for the exponential initial condition in Section~\ref{sec: closing vecs numerics}.

% To further comment on the approximation of a point mass by (\ref{eqn:aafba}), recall the . We found that this closing operator has difficulty approximating mass at .

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/l1_cdf_error_formatted.pdf}
	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the exponential initial condition. The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
	\label{fig: hitting time exp} 
\end{figure}
\begin{figure}[h]
	\centering 
	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/pdf_order21.pdf}%
	\caption{Approximations of the PDF of the first hitting time for Model~\ref{model: simple} with the exponential initial condition. The blue lines were obtained from the dimension-21 DG scheme, the purple lines were obtained from the dimension-21 QBD-RAP scheme and the gold lines were obtained from the dimension 22 DG-lin-lim scheme. The DG scheme displays oscillations around the discontinuities at \(t=1\). The QBD-RAP scheme has an oscillation near \(t=0\). } 
	\label{fig: hitting time oscillation exp} 
\end{figure}
\exampleFloatBarrier
\paragraph{Point mass initial condition}
Figure~\ref{fig: hitting time pm} shows the KS error (left) and \(L^1\) error for the CDF (right) for the five numerical approximation schemes applied to the hitting time problem with the point mass initial condition. The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation as we might expect given the discontinuity. Of the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes all appear to converge, with the QBD-RAP converging fastest and the DG-lin-lim and uniformisation schemes performing similarly. The DG scheme appears to perform best, however, oscillations are present in the DG approximation. In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis, and for the DG-lin-lim scheme with dimension 22. Observing Figure~\ref{fig: hitting time oscillation} the DG approximation displays oscillations around and the discontinuity at \(t=1\) while the other two schemes do not. Of the two positivity preserving schemes in Figure~\ref{fig: hitting time oscillation}, the QBD-RAP scheme appears to approximate the discontinuity best. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/l1_cdf_error_formatted.pdf}
	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition. The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: hitting time pm} 
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/cdf_order21DG_and_sims.pdf}%
	\caption{Approximations of the CDF of the first hitting time in Phase~1 for Model~\ref{model: simple} with the point mass initial condition. The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dashed line from the dimension 22 DG-lin-lim scheme. The empirical CDF obtained via simulation is plotted as the teal dashed line. The DG scheme displays oscillations. } 
	\label{fig: hitting time oscillation} 
\end{figure}

In summary, the results of this experiment are somewhat similar to the last experiment on transient distributions; the DG-lim scheme performs poorly, the DG scheme exhibits oscillations, of the three other positivity preserving schemes, the uniformisation scheme performs poorest, while the QBD-RAP and DG-lin-lim schemes perform similarly, with the QBD-RAP scheme performing slightly better for the point mass initial condition and the DG-lin-lim scheme performing slightly better for the exponential initial condition.

\FloatBarrier
\section{First-return distributions of fluid-fluid queues}\label{sec: ffq num}
Now consider the fluid-fluid queue first analysed in \cite{blnos2022} which is a modified version of a model first presented in \cite{lnp13}. We refer the reader to \cite{blnos2022} for more on the DG scheme applied to this model. 
\begin{model}\label{model: ffq}
	Consider a stochastic fluid-fluid queue $\{(Y(t),X(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{Y(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$. Both $\{X(t)\}$ and $\{Y(t)\}$ have a regulated boundary at 0. Here, the state $11$ indicates inputs to both buffers being \textnormal{\textsc{on}}, the state $00$ indicates both being \textnormal{\textsc{off}}, the state $10$ is when only the first input is \textnormal{\textsc{on}}, and the state $01$ is when only the second is \textnormal{\textsc{on}}. The input of Buffer~$k$ is switched from \textnormal{\textsc{on}} to \textnormal{\textsc{off}} with rate $\gamma_k$, and from \textnormal{\textsc{off}} to \textnormal{\textsc{on}} with rate $\beta_k$, for $k = 1, 2$. Thus, the infinitesimal generator $T$ for $\varphi(t)$ is given by 
	\begin{align*} 
		T = \left[ \begin{array}{cccc} -(\gamma_1 + \gamma_2) & \gamma_2 & \gamma_1 & 0 \\
							\beta_2 & -(\gamma_1 + \beta_2) & 0 & \gamma_1 \\
							\beta_1 & 0 & -(\gamma_2 + \beta_1) & \gamma_2 \\
							0 & \beta_1 &\beta_2 &-(\beta_1 + \beta_2)
	\end{array}\right].
	\end{align*}

	The net rates of change for $X(t)$, denoted $c_i$, are given by 
	% 
	\begin{align*} 
	& (c_{11},c_{10},c_{01},c_{00})   =  \begin{array}{rrrr} 
	(\lambda_1-\theta_1,  & \lambda_1 -\theta_1, &  -\theta_1, & -\theta_1),
	\end{array}  
	\end{align*} 
	% 
	and the net rates of change for $Y(t)$, denoted $r_i$, are as follows  
	%  
	\begin{align*} 
	(r_{11},r_{10},r_{01},r_{00})  & = \left\{ \begin{array}{lrrrll}  
	% (\lambda_2 -\kappa, & \;\;\;\;0, & \lambda_2 - \kappa, &  \;\;\;\;0) & \text{if } X_t = 0,  Y_t = 0,\\
	%  \vspace*{-0.3cm} \\
												(\lambda_2 - \kappa, & \;\;\;-\kappa,  & \lambda_2 - \kappa, & -\kappa) &\text{if } X_t = 0, \\
												\vspace*{-0.3cm} \\
	%	(\lambda_2 -\theta_2,  &  0,  & \lambda_2 - \theta_2,  & \;\;\;\; 0) & \text{if } X_t \in (0,x^*), Y_t = 0,\\
	%											  \vspace*{-0.3cm} \\
		(\lambda_2 - \theta_2, & -\theta_2, & \lambda_2 - \theta_2, & -\theta_2) & \text{if } X_t \in (0,x^*),\\
												\vspace*{-0.3cm} \\
		(\;\;\;\;\;\;\; \lambda_2, & \;\;\;\;0, & \lambda_2, &  \;\;\;\;0) & \text{if } X_t \geq x^*. \end{array}
												\right.
	\end{align*}

	For our numerical experiments, we use the parameter choices given in~\citep{lnp13}: 
	% 
		\begin{align} 
			\label{eqn:parameters}
		\gamma_1 & =11, \quad  \beta_1 = 1, \quad \lambda_1 = 12.48, \quad  \theta_1 = 1.6, \quad  \kappa = 2.6, \\
			\label{eqn:parameters-2}
		\gamma_2 & = 22, \quad \beta_2  = 1, \quad  \lambda_2 = 16.25, \quad \theta_2 = 1.0, \quad x^* = 1.6.
		\end{align} 
	We consider the initial distribution which is a point mass at \(Y(0)=0,\, X(0)=5,\, \varphi(0)=01\). 
\end{model}

While the true problem has an unbounded domain $[0,\infty)$, the approximations require the domain of approximation to be a finite interval. Here we choose an upper bound of \(b=48\) and place a regulated boundary at the upper boundary. The effect of this truncation can be partly quantified by evaluating \(\lim\limits_{t\to\infty}\mathbb P\left(X(t) > 48\right)\approx 5.83 \times 10^{-9}\).

We obtained approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes, all of which used a cell width of \(\Delta=0.4\). The approximate generator matrix resulting from the approximation is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}. Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes the approximations to \(\mathbb \Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively (see \cite[Chapter~7]{p2019} and \cite{bgnp2021} for details on the latter). For the DG scheme the approximation of \(\mathbb \Psi(s)\) is not as well-understood.\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}

Ultimately, we approximate the first-return distribution 
\begin{align}\label{eqn: first return Y 1}
	\mathbb P(X(\zeta_{Y}(\{0\}))\leq x, \varphi(\zeta_{Y}(\{0\}))=i\mid \bs X(0)\sim \mu),
\end{align}
where we recall \(\zeta_{Y}(E) = \inf\{t>0\mid Y(t)\in E\}\) is the first hitting time of \(\{Y(t)\}\) on the set \(E\). For Model~\ref{model: ffq}, it is only possible for the process \(Y(t)\) to return to \(0\) at time \(t\) when \((X(t),\varphi(t))\in[0,1.6)\times \{10,00\}\), so we evaluate the approximations over this region only. We use a grid of 10,001 points on which we evaluate the approximations of the CDF in each phase.

For a `ground truth' comparison, we simulated 5,000,000 realisations of the fluid-fluid queue and recorded the value of \(\{X(t)\}\) and \(\{\varphi(t)\}\) at the time of first return of the second fluid, \(\{Y(t)\}\). The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations was computed. To account for Monte-Carlo errors we used a bootstrap with 1,000 bootstrap samples to construct 1,000 bootstrap samples of the error estimates and ultimately estimated the 5th and 95th percentiles of the error distribution. Each of the 1,000 bootstrap samples was constructed by resampling the original 5,000,000 realisations 5,000,000 times with replacement.

In Figures~\ref{fig: ffq return cts} and~\ref{fig: ffq cell probs} we plot the error metrics for the approximations to the distribution~(\ref{eqn: first return Y 1}). The DG scheme performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors. The QBD-RAP scheme is second best and the uniformisation scheme appears to be the slowest to converge. Here, the first return distribution appears to be smooth, hence we might expect the DG scheme to perform well.

Note that there is a significant number other sources of error here; machine precision errors, errors in solving the matrix Riccati equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors. Furthermore, for the QBD-RAP, since the parameters \(\bs \alpha,\,\bs S,\,\bs s\,\) and \(\bs D\) are found numerically, then there is another source of error from this. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cdf_error_formatted.pdf}
	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}. The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: ffq return cts} 
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cell_probs_error_formatted.pdf}%
	\caption{Cell-wise errors between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}. The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: ffq cell probs} 
\end{figure}
\exampleFloatBarrier

By modifying slightly Model~\ref{model: ffq}, we can construct a first return distribution which is discontinuous. 
\begin{model}\label{model: ffq2}
	Consider a fluid-fluid queue which is the same as Model~\ref{model: ffq} except 
	\begin{align}
		r_{00}(X(t)) = \begin{cases}
			-\kappa, & \mbox{ if }X(t)=0,\\
			-\theta_2, & \mbox{ if }X(t)=\in(0,x^*),\\
			\theta_2, & \mbox{ if }X(t)\geq x^*,
		\end{cases}
	\end{align}
	with an initial distribution which is a point-mass at \(Y(0)=0,\, X(0)=2, \varphi(0)=00\). 
\end{model}
As before, we use the DG, uniformisation and QBD-RAP schemes to approximate the model, and compare the approximations to simulations.

For Model~\ref{model: ffq2} there is a point mass at \(X(\zeta_{Y}(\{0\}))=1.2\) of magnitude \(e^{-(\beta_1+\beta_2)}\times 0.5\), which occurs when the phase remains in \(\varphi(t)=00\) for all \(t\in [0,\zeta_{Y}(\{0\})]\).

Figures~\ref{fig: ffq return discts} and~\ref{fig: ffq2 cell probs} plot the error metrics for the first return distribution. Observing Figure~\ref{fig: ffq return discts} we see that the DG scheme performs well, followed by the QBD-RAP scheme, then the uniformisation scheme performs worst. However, the DG scheme results in an oscillatory solution with the CDF taking impossible values (decreasing at points) as shown in Figure~\ref{fig: ffq2 oscillation}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cdf_error_formatted.pdf}
	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}. The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: ffq return discts} 
\end{figure} 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cell_probs_error_formatted.pdf}%
	\caption{Cell-wise error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing in each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq2}. The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes. Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines. The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.} 
	\label{fig: ffq2 cell probs} 
\end{figure}
\begin{figure}[h] 
	\centering
	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/ffq/discts/phase_4_cdf.pdf}%
	\caption{Approximations of the CDF of the distribution of \(X\) at the time \(\zeta_Y(\{0\})\) in Phase~1 for Model~\ref{model: ffq2}. The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dotted line is the empirical CDF obtained via simulation. The DG scheme displays oscillations around the discontinuity. } 
	\label{fig: ffq2 oscillation} 
\end{figure}

In summary, for the smooth problem the DG scheme was highly effective, significantly outperforming the positivity preserving QBD-RAP and uniformisation methods. For the discontinuous problem, the DG scheme displayed oscillations, but produced the lowest errors. For both problems, of the positivity preserving methods (QBD-RAP and uniformisation scheme), the QBD-RAP scheme performed the best.

\FloatBarrier

\section{Discussion}
In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing schemes; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme. In general, the numerical experiments that the smoother the problem is the better the performance of the DG scheme, and it emphatically outperforms the other two schemes for smooth problems. However, for problems with discontinuities the DG approximation can exhibit oscillations, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.

To avoid the problems of oscillations we can sometimes employ a \emph{slope limiter} with the DG scheme, which effectively reduces the scheme to linear in the regions where oscillations are detected. In some contexts, we implemented two slope-limited DG schemes, the DG-lim scheme which takes a high order DG scheme and limits the solution as necessary, and the DG-lin-lim, which is a linear approximation on a finer grid designed to use approximately the same computational resources as the other schemes we consider. The numerical experiments demonstrate a significant loss of accuracy in the approximation when a DG-lim scheme is used for discontinuous problems. The numerical experiments suggest that the DG-lin-lim scheme can perform well, and is similar to the performance of the QBD-RAP scheme, in the presence of discontinuities. For highly discontinuous problems (i.e.~problems with point masses) the QBD-RAP scheme can outperform the DG-lin-lim scheme. %For the application of the schemes to fluid-fluid queues, there is no obvious way to apply the concept of a slope limiter. %When the slope limiter detects oscillations in the approximate solution, it reduces the DG scheme to linear in the region surrounding the oscillations, otherwise, the slope limiter leaves the solution unchanged. Thus, the slope limiter permits high-order approximations away from oscillations, while also removing oscillations.

As a first step in the numerical experiments, in Section~\ref{sec: recon num}, we examined the ability of each scheme to approximate different initial conditions. For the DG scheme, this is equivalent to a projection of the initial condition on to a set of basis polynomials, for the uniformisation scheme this is equivalent to projecting the initial condition on to a basis of constant functions, and for the QBD-RAP scheme the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}. Section~\ref{sec: comp} demonstrates that the DG scheme (projection) can result in oscillations and negative regions in the approximation when the initial condition is discontinuous. The uniformisation and QBD-RAP schemes avoid this problem, but appear to have higher errors and the QBD-RAP scheme appears to have the largest errors. For discontinuous initial conditions the rates of convergence are comparable for all three schemes. However, when the initial condition to be approximated is sufficiently smooth, then the DG approximation is superior. The QBD-RAP scheme performed the poorest for the examples considered.

Next, in Section~\ref{sec: wave num}, we instrumented the performance of approximations for a simple travelling wave model with various initial conditions. For this model the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\). This model is useful as it allows us to instrument the ability of the schemes to capture the flow of probability without stochastic dynamics, and since the solution is known, so there is no need for simulation. With the travelling wave model we demonstrate that, for problems with discontinuities, the DG scheme can display oscillations, while the other schemes (DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes) avoid oscillations. For discontinuous problems the rates of convergence of the QBD-RAP and DG-lin-lim schemes is similar. Interestingly, even though the QBD-RAP scheme performed worst for approximating initial conditions in the previous section, it outperformed the uniformisation scheme for this model, demonstrating that the QBD-RAP scheme can capture the dynamics of the flow of probability better than the uniformisation scheme. For smooth problems the DG scheme is superior.

We then instrumented the performance of the approximations on a simple fluid queue with two phases. First, in Section~\ref{sec:stat}, we looked at the stationary distribution, which is known to be smooth. Since the problem is smooth, then the DG scheme was superior as expected. Of the uniformisation and QBD-RAP schemes, the QBD-RAP scheme gives more accurate solutions.

In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same model and consider two different initial conditions, a point-mass and an exponential initial condition. The discontinuous initial condition results in a discontinuous transient distribution. As for the exponential initial condition, this example demonstrates that, even if the initial condition appears `nice', it can still result in non-smooth (i.e.~non-differetiable) solutions. The numerical evidence suggests that the DG scheme can display oscillations, while the other schemes do not. The DG-lim scheme detects the oscillations and reduces the approximation to linear. Of the uniformisation, DG-lin-lim and QBD-RAP schemes, the latter two performs similarly and better than the uniformisation scheme. Of the DG-lin-lim and QBD-RAP schemes the QBD-RAP scheme performed better in the presence of a point mass while the DG-lin-lim scheme performed better with the exponential initial condition.

Next, Section~\ref{sec: return approx} looked at hitting times for the same fluid queue with two initial conditions, an exponential initial condition and a point-mass. We looked at the hitting time of the fluid level on the points \(\{0\}\) and \(\{1\}\). For this problem there is never any in-flow of mass at the boundaries of the interval and so, for a solution to be continuous, the initial condition needs to be chosen carefully, otherwise discontinuities in the transient distribution may result, as is the case for both initial conditions here. The numerical results suggest that, due to the discontinuities in the problems, the DG scheme may display oscillations. Since the uniformisation, DG-lin-lim, and QBD-RAP schemes can handle discontinuities, they perform suitably, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation.

Lastly, we applied the DG, uniformisation and QBD-RAP schemes to two simple fluid-fluid queues in Section~\ref{sec: ffq num}. In the first fluid-fluid queue, which appears to have a smooth solution, the DG scheme performs very well. Of the two positivity preserving schemes the QBD-RAP scheme performs better than the uniformisation scheme. In the second example, which has a discontinuity, the DG scheme produces the lowest errors, but exhibits oscillations in the solution. The QBD-RAP and uniformisation schemes do not produce oscillations and, of the two, the QBD-RAP scheme performs best.

In conclusion, when the problem is known to be smooth, the DG scheme is very likely to produce excellent results. However, for discontinuous problems, the scheme can show oscillations and infeasible or negative regions of the solution. The slope limiter overcomes this, but reduces the accuracy of the DG scheme to linear near discontinuities, sometimes severely affecting the quality of the approximation. A piecewise-linear DG scheme with a limiter, such as the DG-lin-lim scheme considered here, can produce satisfactory results. The uniformisation and QBD-RAP schemes are other alternative approximation schemes which avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF and have a stochastic interpretation. Of the uniformisation and QBD-RAP schemes, the latter often produced lower errors and performed similarly to the DG-lin-lim scheme. Moreover, the uniformisation and QBD-RAP schemes have interpretations which can aid in their analysis (such as justifying the Riccati equation to approximate \(\mathbb \Psi\) for a fluid-fluid queue). The uniformisation and QBD-RAP schemes also result in linear approximations, whereas the DG-lim and DG-lin-lim schemes do not and this can be beneficial in certain contexts.