### 2022-02-17 front/acknowledgements.tex

%!TEX root = ../thesis.tex
\chapter{Acknowledgements}
\label{ch:acknowledgements}

First and foremost, to my principle supervisor, Nigel Bean. Thanks for all the support, the interest you have taken in this work, and the faith you have shown in me allowing me to essentially do whatever I want for this thesis. The effort you have put in to supervising me has been much appreciated.

To Peter Taylor, my co-supervisor, thanks for joining me on this quest. Thanks for the insightful discussions on maths, cricket and all the other important happenings of the world during our meetings. You've provided an invaluable perspective on my work and on mathematics as a science.

To Giang Nguyen. You were the one who initially inspired me to take up mathematics research which led me to where I am now. You gave me so much of your time and went well above and beyond to support me. Thank you.

To my friend Caitlin, thanks for all the support, coffee, and company, over the last five years. It has been so enjoyable. I will miss it.

To my friends Dennis, Phil and Andrew. I can't understate the value of this group (and Caitlin too). I've been lucky to be able to share ideas, get advice, and receive mentorship and support from all of you.

To my wife, Alice. Thank you for your support over the last five years. I've learnt so much about life from you.