### 2022-02-17 intro/chapter.tex

%!TEX root = ../thesis.tex
\chapter{Introduction\label{ch:intro}} 
% intro to intro
% contextual background
	% Fluid queues 
		% definition
		% applications 
		% existing results 
	% Fluid-fluid queues
		% definition 
		% basic analysis in terms of generators
		% the need for approximation and how the old work doesn't fit here
	% The DG method
		% pertition into cells
		% project on to a basis
		% problems: oscillations
		% soutions and why they dont fit
		% observation about constant-basis methods
		% this thesis asks the question: motivated by the constant basis being a probability model, can we dream up a more accurate probability model to approximate a fluid queue?
	% On the structure of the constant basis/uniformisation method (QBD)
	% least variable is PH, so what about MEs?
% structure of the thesis
	% rest of chapter 1 gives mathematical preliminaries
	% chapter 2 explains the DG method, problems/oscillations, slope limiting, and appllication to SFFMs
	% inspired by the structure of the order-1 scheme, and to solve the negativity problem, chapter 3 introduces a new approximation method; The QBD-RAP. In this chapter we derive the intuitively of the model and explain it's dynamics.
	% we then move to convergence, with chapter 4 proving a convergence of the QBD-RAP up to the first hitting time of the fluid queue on the edges of an interval. technical results and extensions are left to the appendix: certain properties of closing vectors, convergence without ephemeral phases, and certain matrix algebraic manipulations. 
	% chapter 5 stitches together the results of chapter 4 to prove a global result about convergence
	% chapter 6 investigate, numerically, the performance of the DG, DG with limiter, uniformisation and QBD-RAP schemes. 
	% chapter 7  makes concluding remarks.
% mathematical preliminaries 
	% CTMCs
	% fluid queues
	% fluid-fluid queues and operator-analytic expressions (all of them)
		% the equations of bo2014
		% meets the partition of the DG/QBD-RAP, and reconstructing the Fil sets 
    % why we cant solve the equations, what we need to approximate to solve it, i.e. B, R, then D, Psi
	% projections
	% phase-type distributions 
		% definition
		% least variable property
	% matrix exponentials
		% definition 
		% properties 
		% verification that parameters give an ME
		% CMEs
	% QBD-RAPs, then QBDs as a special case
		% orbit processes, their interpretation for PH and how they differ for MEs
	% convergence theorems 
		% DCT 
		% Continuity of Laplace transforms

% THE ROUGH IDEA & CONTEXT
% 	THE CONTEXT: FLUID-FLUID QUEUES
%	PDEs 
%	NON-NEGATIVITY & CONSERVATION 
% MATHEMATICAL PRELIMINARIES

A fluid queue is a two-dimensional stochastic process \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\). The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC) and determines the rate at which \(\{ X(t)\}\) moves. The level process, \(\{ X(t)\}_{t\geq0}\), is real-valued, continuous, piecewise linear and deterministic given \(\{\varphi(t)\}\).

Typically, the phase, \(\varphi(t)\), represents the underlying operating state or environment of a system at time \(t\). The level variable, \(X(t)\), describes some continuous variable of the system. Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982}, a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005}, environmental modelling \citep{wurm2020}, dam storage \citep{loynes1962}, and more. For example, \cite{anick1982} model a \emph{data handling switch} in which there are \(N\) information sources which asynchronously switch from ON to OFF, or OFF to ON. The phase process \(\varphi(t)\) models the number of data sources which are ON at time \(t\), and the rate at which data is received when the buffer is on is assumed (without loss of generality) to be 1. The switch stores data in a buffer and processes the data at rate \(c\), thus the net rate of change of the buffer at time \(t\) when there are \(r\) information sources which are ON is \(r-c\). The level process, \(X(t)\), models the amount of data in the buffer at time \(t\) and the rate at which \(X(t)\) changes which \(\varphi(t)=r\) is \(c_r=r-c\). The model described above could be used to aid in the analysis and design of the data handling switch, for example, by determining the minimum rate \(r\) for which the system is \emph{stable} in steady state, which practically means that the processing power of the switch is large enough that the amount of data in the buffer does not grow without bound.

The success of fluid queues largely lies in their mix of flexibility, any (finite) number of phases can be used, and their parsimony and analytic tractability which enables effective transient and stationary analysis \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018,bnp2018}. Fluid queues are relatively well studied. Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods e.g.~\cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \cite{anick1982,kk1995,blnos2022}. %For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest. Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method. Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.

More recently, \cite{bo2014} extends fluid queues to so-called \emph{stochastic fluid-fluid queues}. In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{Y(t)\}\) moves. The addition of the second level process \(\{Y(t)\}\) greatly increases the modelling potential. For example, \cite{hydro,bo2014} describe a model for a hydroelectric power generator which can be operated `on design' and `off design'. When the system operates `off design' there system is less efficient and there is more wear, but may be used to optimise overall system performance. Further, maintenance of the generator takes place periodically to improve its performance and lifespan. The system can plausibly be modelled as a fluid queue, \(\{(\varphi(t),X(t))\}\), where the state space of \(\{\varphi(t)\}\) consists of states representing the systems states `start', `stop', `on design', `off design', `maintenance', and `idle'. The level variable \(\{X(t)\}\) take values in \([0,1]\) and represents the deterioration level of the system with 0 being perfect working order and 1 meaning the system needs replacement. The second level process \(\{Y(t)\}\) could be used to model the revenue of the power generator, which, logically, should depend on the state of the system, \(\{\varphi(t)\}\). Moreover, the revenue could also depend on the deterioration level \(\{X(t)\}\), as, for example, when the system is brand new it may operate more efficiently, and when it is near broken it may be less efficient.

% In this paper, we are interested in the following modeling potential. Suppose that, besides
% the level X (t) at time t, we would like to model some other continuous performance measure
% Y (t), such as the net profit at time t for example. To illustrate this, consider a hydro-power
% generator, which can be operated “on design” and “off design”. The latter causes increased
% wear, and is less efficient, but can be useful to optimize overall system operation. The generator
% must be periodically maintained, in order to improve its performance and prolong its lifespan.
% An important related problem is the evaluation of maintenance strategies and the impact of
% [1–6,10,12–16,28,29][7,12,21]
% N.G. Bean, M.M. O’Reilly / Stochastic Processes and their Applications 124 (2014) 1741–1772 1743
% maintenance on reliability. This problem can be modeled as a bounded SFM {(φ(t), X (t)), t ≥ 0}
% with S consisting of operating phases (start, stop, on design, off design, maintenance, idle) and
% deterioration level X (t) ∈ [0, 1], where 0 means the generator is brand new and 1 means it needs
% replacement, and it is assumed that once level 1 is reached, the process {(φ(t), X (t)), t ≥ 0}
% restarts itself from level 0 [12].
% Now, in order to describe the continuous revenue process {Y (t), t ≥ 0} associated with
% the process {(φ(t), X (t)), t ≥ 0}, it may be plausible to use a SFM driven by {φ(t), t ≥ 0}.
% This is because the rate at which the revenue is increasing/decreasing at time t, ideally should
% depend on the phase φ(t). However, for practical reasons, we may additionally require that the
% revenue Y (t) at time t depends on the operating level X (t). For example, if the generator is
% newer, it may operate more efficiently, produce more energy and require less-costly maintenance.
% Hence, the rates at which the revenue level Y (t) is increasing/decreasing may depend both on
% the phase i ∈ S and the operating level X (t) = x at time t, and so to reflect this, we denote
% these rates by ri (x). Summarizing this, we propose a new class of fluid models, denoted by
% {(φ(t), X (t), Y (t)), t ≥ 0}. We refer to this class as the stochastic fluid–fluid model, since it is
% a stochastic fluid model with level Y (t) driven by {(φ(t), X (t)), t ≥ 0}, which is also a stochastic
% fluid model with level X (t). The buffers collecting fluid from the process {X (t), t ≥ 0} and
% {Y (t), t ≥ 0} are referred to as X and Y , respectively.

The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential, however comes at some cost to the ease of their analysis. The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisation have been suggested. Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation scheme throughout this thesis. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where cells are all have equal width, the operator is projected onto a basis of piecewise constant functions, and an upwind flux is used to approximate the flow of mass between cells. This is also equivalent to a finite-volume scheme.

%A QBD can be viewed as a two-dimensional CTMC, \(\{(L(t),\varphi(t))\}_{t\geq0}\), where \(\{L(t)\}\) is the discrete level process, and \(\{\varphi(t)\}_{t\geq0}\) is the phase process. The level process \(\{L(t)\}\) is skip free, meaning that, given the process is at level \(L(t)=\ell\), is may only jump to \(\ell+1\) or \(\ell-1\) at jump epochs. The sojourn time of \(\{L(t)\}\) in a given level follows a phase-type distribution

One drawback of the uniformisation scheme is that convergence can be relatively slow compared to higher-order DG schemes. However, in the context of approximating fluid queues, one advantage of the uniformisation scheme and, equivalently, a DG scheme with constant basis functions, is that the uniformisation scheme guarantees probabilities computed from the approximation are positive \citep[Section~3.3]{koltai2011}. One justification for the positivity preserving property of the uniformisation scheme is from its interpretation as a stochastic process. For higher order DG schemes there is no such interpretation and positivity is not guaranteed \citep[Section~3.3]{koltai2011}. Moreover, higher-order DG approximation schemes may produce negative and oscillatory solutions, particularly when discontinuities or steep gradients are present. This is problematic for probabilistic problems as we know that probabilities must be positive. Methods to navigate the problem of negative and oscillatory solutions have been developed, such as filtering and slope limiting (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6 and references therein). Slope limiting alters the discretised operator in regions where oscillations are detected and reduces the order of the approximation to linear in these regions but ensures the solution will be non-negative. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.

Depending on the context, filtering of the approximate solution to remove oscillations may not necessarily guarantee a strictly non-negative approximation, or may result in severe smearing of the solution at discontinuities or regions with steep gradients \citep[Section~5.6.1]{nodalDGBook}. Moreover, filtering requires us to make a trade-off between filtering enough of the oscillations away while maintaining sufficient accuracy -- a choice which may not be obvious \emph{a priori}. Slope limiting does guarantee positivity but reduces the approximation to linear where oscillations in the approximate solution are detected \citep[Section~5.6.1]{nodalDGBook}. Furthermore, limiting and filtering do not distinguish between natural oscillations, which are a fundamental feature of the solution and spurious oscillations which are caused by the approximation scheme, and they may remove both from the approximation. This can lead to an unnecessary loss of accuracy in the approximation (see \citep{nodalDGBook},~Example~5.8). Other drawbacks of these methods is that they result in non-linear operators, so we cannot use matrix algebra for computations, and they incur an additional computational cost on top of the DG scheme.

% In the context of approximating the first return operator of a fluid-fluid queue, the application of a slope limiter is not possible since we need the approximated operators to be linear.

In approximating the first return operator of a fluid-fluid queue we first approximate an operator-Riccati equation by a matrix-Riccati equation by substituting in matrix approximations to the infinitesimal generator of the fluid queue which are constructed via the DG scheme \citep{blnos2022}. We then solve the matrix-Riccati equation via an iterative procedure \citep{bean2005b,blnos2022}. Since the operators constructed with limiters are non-linear (not matrices) then we cannot apply limiters in this case. Alternatively, we could apply a post-processing based on slope limiters, or filtering. Both of these post-processing methods incur a computational cost and need to be applied on a case-by-case basis depending on the initial condition.

Furthermore, for the DG method there is a limited theoretical backing as to why we may approximate the operator-Riccati equation with the matrix-Riccati equation other than it is a sensible thing to do, and it seems to work in practice. In the case of the QBD approximation of \cite{bo2013} the approximate matrix-Riccati is more justified as in it can be derived by considering the first-return probabilities for a fluid queue driven by the approximating QBD where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\) \citep{bean2005}.

Hence, for the approximation of fluid queues with application to fluid-fluid queues, on the one hand we can use the DG scheme which can produce high-order approximations when the problem is smooth, but can display spurious oscillations and result in negative probability estimates in the presence of discontinuities. While, on the other hand, we can use the uniformisation scheme which constructs in a linear operator to approximate the generator and guarantees non-negative probability estimates, but convergence is relatively slower.

Motivated by this, this thesis derives a new approximation to a fluid queue. The construction of the approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to approximate the sojourn time of the fluid level in a given interval on the event that the phase of the fluid is constant. The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution, so in this sense, the best approximation to this deterministic sojourn time. Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation. Recently, there has been much work on a class of concentrated matrix exponential distributions \citep{hhat2020} which are postulated to be the least-variable matrix exponential distributions of a given order. Matrix exponential distributions generalise Phase-type distributions; they have the same functional form, without the restriction that the distribution has an interpretation in terms of the absorption time of a continuous-time Markov chain. A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \citep{bn2010}, extends QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times. Thus, by using matrix exponential distributions, we attempt to construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}, while retaining a stochastic interpretation.

As the QBD-RAP has a stochastic interpretation then the approximations it produces are guaranteed to have non-negative density functions. Further, for the QBD-RAP (and uniformisation) scheme non-negativity is guaranteed for any initial condition without any further computation or post-processing. For the approximation of fluid-fluid queues, the matrix-Riccati equation we solve to approximate the first-return distribution of a fluid-fluid queue can be derived by considering the first-return probabilities for a RAP-modulated fluid queue driven by the QBD-RAP \citep{p2019,bgnp2021}. The stochastic interpretation can also be leveraged to aid in the analysis of the approximation. Another attractive property is that the approximations to operators obtained via the QBD-RAP (and uniformisation) scheme are linear which is necessary for the application to approximating the first-return operator of fluid-fluid queues and may also provide other computational and analytical benefits.

The structure of this thesis is as follows. The next chapter, Chapter~\ref{ch:math}, is dedicated to mathematical preliminaries and introduces the main mathematical objects and tools which we will need. In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way that makes clear exactly how the approximations we develop correspond to the theoretical operators. Chapter~\ref{ch:math} also introduces the existing literature and gives context to this thesis.

Chapter~\ref{ch:galerkin} demonstrates a way that we can use the discontinuous Galerkin (DG) method to approximate certain operators and distributions of fluid and fluid-fluid queues. Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:galerkin} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and negativity. At the end of Chapter~\ref{ch:galerkin} we also introduce other numerical procedures which we use in our numerical investigations in Chapter~\ref{sec: numerics}.

In the following chapter, Chapter~\ref{sec: construction and modelling}, we develop a new QBD-RAP approximation scheme which, due to its interpretation as a stochastic process, ensures all estimates of probabilities will be non-negative. The chapter takes a stochastic modelling approach to developing the approximation scheme.

Once we have established the new approximation scheme, we then prove a convergence of the approximation scheme in Chapters~\ref{sec: conv}~and~\ref{ch: global results} via certain Laplace transforms with respect to time. Chapter~\ref{sec: conv} utilises methods and ideas relating specifically to QBD-RAPs and matrix exponential distributions. Ultimately, Chapter~\ref{sec: conv} proves that, on the event that the QBD-RAP has not-yet seen an \emph{orbit restart epoch} (which is much like a change of level, but not always), certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue. This is a local convergence result as it relates only to convergence to the fluid queue in an interval.

Chapter~\ref{ch: global results} then looks to extend the local convergence results of Chapter~\ref{sec: conv} to a global convergence result on the whole domain of approximation. Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on things such as the strong Markov property, time-homogeneity and the Law of total probability. In Chapter~\ref{ch: global results} we first consider the discrete-time process which the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch. We prove that the embedded process converges in distribution to a corresponding embedded process of the fluid queue. The rest of Chapter~\ref{ch: global results} then attempt to prove a global converge result. The main result being Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (weakly with respect to the spatial and temporal variable) to the distribution of the fluid queue.

Once we have established methods for approximating fluid queues, Chapter~\ref{sec: numerics} then numerically explores some properties of the approximations. Given that the QBD-RAP scheme was developed so that it guarantees positivity of the approximation, we largely focus on problems with discontinuities.

Finally, Chapter~\ref{ch: conclusion} makes concluding remarks.

To help the reader understand the notations used in the DG scheme, we provide a small toy model example in Appendix~\ref{appendix:example}. Appendix~\ref{sec:properties} provides a proof that the DG scheme conserves probability. The rest of the appendices contain technical results relating to proving the convergence of the QBD-RAP approximation scheme. Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operators} introduced as part of the QBD-RAP scheme have the properties we claim they do to prove convergence in Chapter~\ref{sec: conv}. Appendix~\ref{app:extend conv} extends some results from Chapter~\ref{sec: conv} to a setting which requires slightly less computation. Lastly, Appendix~\ref{appendix: kronecker} provides some algebraic results which help us to manipulate certain Laplace transforms from Chapter~\ref{sec: conv}.